Hierarchical cortical entrainment orchestrates the multisensory processing of biological motion

Version 1

Abstract

Abstract

When observing others’ behaviors, we continuously integrate their movements with the corresponding sounds to achieve efficient perception and develop adaptive responses. However, how human brains integrate these complex audiovisual cues based on their natural temporal correspondence remains unknown. Using electroencephalogram, we demonstrated that cortical oscillations entrained to hierarchical rhythmic structures in audiovisually congruent human walking movements and footstep sounds. Remarkably, the entrainment effects at different time scales exhibit distinct modes of multisensory integration, i.e., an additive integration effect at a basic-level integration window (step-cycle) and a super-additive multisensory enhancement at a higher-order temporal integration window (gait-cycle). Moreover, only the cortical tracking of higher-order rhythmic structures is specialized for the multisensory integration of human motion signals and correlates with individuals’ autistic traits, suggesting its functional relevance to biological motion perception and social cognition. These findings unveil the multifaceted roles of entrained cortical activity in the multisensory perception of human motion, shedding light on how hierarchical cortical entrainment orchestrates the processing of complex, rhythmic stimuli in natural contexts.

Introduction

The perception of biological motion (BM), the movements of living creatures, is a fundamental ability of the human visual system that is crucial in survival and social situations. Extensive evidence shows that humans can readily perceive BM from a visual display depicting just a handful of light dots attached to the head and major joints of a moving person (Blake & Shiffrar, 2007). Nevertheless, in real life, BM perception often occurs in multisensory situations, e.g., one may simultaneously hear footstep sounds while seeing others walking. The integration of these visual and auditory BM cues based on congruency in stimulus contents or temporal relationships can facilitate the detection and discrimination of BM stimuli (Mendonça et al., 2011; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2013; van der Zwan et al., 2009). Remarkably, such a cross-modal effect appears to engage an audiovisual integration (AVI) mechanism specific to BM, as the effect disappeared when the visual BM signals were deprived of characteristic kinematic cues but not low-level motion attributes through stimulus inversion (Brooks et al., 2007; Thomas & Shiffrar, 2010), and the temporal windows of perceptual audiovisual synchrony are different between BM and other motion stimuli with constant motion speed or gravity-incompatible accelerations (Arrighi et al., 2006; Saygin et al., 2008). Despite the behavioral evidence, the neural basis for the AVI of BM signals based on their natural multisensory correspondence remains largely unclear.

An intrinsic property of human movements (such as walking and running) is that they are rhythmic and accompanied by frequency-congruent sounds. The AVI of such rhythmic stimuli may involve cortical entrainment, a process that the neural oscillations in cortical networks entrain to external rhythms and show increased activity or phase coherence at corresponding frequencies (Bauer et al., 2020; Lakatos et al., 2019). Studies based on simple or discrete stimuli have found that temporal congruency in auditory and visual rhythms significantly enhances the cortical tracking of rhythmic stimulations in both modalities (Nozaradan et al., 2012b). Unlike these stimuli, BM conveys complex hierarchical rhythmic structures that could be extracted from integration windows at different temporal scales. For example, the human locomotion movement has a narrower integration window consisting of each step (i.e., step cycle) and a broader integration window incorporating the opponent motion of the two feet (i.e., gait cycle). A recent study suggests that neural entrainment to these hierarchical kinematic structures contributes to the spatiotemporal integration of visual BM cues in different manners (Shen, Lu, Yuan, et al., 2023). However, it remains open whether and how the cortical tracking of rhythmic signals underpins the AVI of BM information.

To tackle this issue, we recorded electroencephalogram (EEG) signals from participants who viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions in Experiments 1a & 1b (Fig. 1). A greater cortical entrainment effect in the AV condition compared to each unisensory condition will indicate significant multisensory gains. Moreover, contrasting the multisensory response with the summation of the unisensory responses serves to distinguish among sub-additive (AV < A+V), additive (AV = A+V), and super-additive (AV > A+V) modes of multisensory integration (see a review by Stevenson et al., 2014). Experiment 2 further examined to what extent the AVI effect was specific to the multisensory processing of BM by using non-BM (inverted visual stimuli) as a control. Inversion disrupts the unique, gravity-compatible kinematic features of BM but not the rhythmic signals generated by low-level motion cues (Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Wang et al., 2022), thus is expected to interfere with the BM-specific neural responses. Participants perceived the visual stimuli accompanied by temporally congruent or incongruent BM sounds. Comparing the congruency effect in neural responses between the upright and inverted conditions provides a unique opportunity to verify whether the AVI of BM involves a mechanism distinct from that underlies the AVI of non-BM information.

Illustrations of audiovisual stimuli and experimental procedures.

The illustration was based on stimuli with a gait-cycle frequency of 1 Hz. (a) Visual stimuli. The left panel depicts the static schematic of upright and inverted point-light walkers. The right panel shows the keyframes from a gait cycle of the BM sequence. The colors of dots and lines between dots are for illustration only and are not shown in the experiments. (b) Auditory stimuli. The auditory sequences contain periodic impulses of footstep sounds whose peak amplitudes occur around the points when the foot strikes the ground. The duration of two successive impulses defines the gait cycle of footstep sounds, which is temporally congruent (Con) or incongruent (InC) with the visual stimuli. (c) Experimental procedure. The color of the visual stimuli changed one or two times within 6 s in the catch trials but did not change in the experimental trials. Participants were required to report the number of changes when the point-light stimulus was replaced by a red fixation. In Experiment 1, participants viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions. The visual stimulus was the BM sequence in the V and AV conditions but a static frame from the sequence in the A condition. Experiment 2 included only the AV condition with different stimulus orientations (upright vs. inverted) and audiovisual congruency (congruent vs. incongruent).

It is also worthy to note that the abilities to process BM information and integrate multisensory inputs vary across individuals and are diminished in populations with autism spectrum disorder (ASD) or even high autistic traits (Feldman et al., 2018; Pavlova, 2012; Wang et al., 2018). Specifically, ASD individuals showed reduced orienting to audiovisually synchronized BM stimuli (Klin et al., 2009), and such impairment in 10-month infancy can predict autism diagnosis at 3 years of age (Falck-Ytter et al., 2018). These findings suggest a possible link between compromised audiovisual BM processing ability and higher autistic-like traits, given that social cognitive deficits in ASD lie on a continuum extending from the clinical to nonclinical populations with different levels of autistic traits (Baron-Cohen et al., 2001). Therefore, here we examined the potential relationship between participants’ neural responses to synchronous audiovisual BM signals and their autistic traits in Experiment 2.

Results

In all experiments, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed one or two times throughout the trial, and there was no color change in other trials. Participants were required to detect the color change of visual stimuli (0-2 times during one trial) to maintain attention. Behavioral analysis on all trials showed that their performances for the task were generally high and equally well in all conditions of Experiment 1a (mean accuracy > 98%; F (2, 46) = 0.814, p = .450, = 0.034), Experiment 1b (mean accuracy > 98%; F (2, 46) = 0.615, p = .545, = 0.026), and Experiment 2 (mean accuracy > 98%; F (3, 69) = 0.493, p = .688, = 0.021), indicating comparable attention state across conditions. The catch trials were excluded from the following EEG analysis.

Cortical tracking of rhythmic structures in audiovisual BM reveals AVI

Experiment 1a

In Experiment 1a, we examined the cortical tracking of rhythmic BM information under V, A, and AV conditions. We were interested in two critical rhythmic structures in the walking motion sequence, i.e., the gait cycle and the step cycle (Fig. 1a). During walking, each step of the left or right foot occurs alternatively to form a step cycle, and the antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, the frequency of a full gait cycle is 1 Hz, and the step-cycle frequency is 2 Hz. The strength of the cortical tracking effect was quantified by the amplitude peaks emerging from the EEG spectra at these frequencies.

As shown in the grand average amplitude spectra (Fig. 2a), both the responses in three conditions showed clear peaks at step-cycle frequency (2 Hz; V: t (23) = 6.964, p < 0.001; A: t (23) = 6.073, p < .001; AV: t (23) = 7.054, p < 0.001; FDR corrected). In contrast, at gait-cycle frequency (1 Hz), only the response to AV stimulation showed significant peaks (V: t (23) = −2.072, p = 0.975; A: t (23) = −0.054, p = 0.521; AV: t (23) = 4.059, p < 0.001; FDR corrected). Besides, we also observed a significant peak at 4 Hz in all three conditions (ps < 0.001, FDR corrected), which might be the harmonic of 2 Hz (see the results of harmonics in Supplementary Information).

Cortical tracking of visual (V), auditory (A), and audiovisual (AV) BM signals at gait-cycle and step-cycle frequencies.

(a) & (d) The amplitude spectra of EEG response in three conditions in Experiment 1a and Experiment 1b, respectively. The solid lines show the grand average amplitude over all electrodes and subjects. The shaded regions depict standard errors of the group mean. Asterisks indicate significant spectra peaks (one-sample t-test against zero; p < .05, FDR corrected). (b) & (e) The normalized amplitude at gait-cycle frequency in the AV condition exceeded the arithmetical sum of those in V and A conditions (AV > A+V), (c) & (f) but the normalized amplitude at step-cycle frequency in the AV condition was comparable to the sum of V and A (AV = A+V). Colored dots represent individual data in each condition. Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; m.s.: .05< p < .10; n.s.: p > .05.

Furthermore, we directly compared the cortical tracking effects between different conditions via a two-tailed paired t-test. At both 1 Hz (Fig. 2b) and 2 Hz (Fig. 2c), the amplitude in the AV condition was greater than that in the V condition (1 Hz: t (23) = 4.664, p < 0.001, Cohen’s d = 0.952; 2 Hz: t (23) = 5.132, p < 0.001, Cohen’s d = 1.048) and the A condition (1 Hz: t (23) = 2.391, p = 0.025, Cohen’s d = 0.488; 2 Hz: t (23) = 3.808, p < 0.001, Cohen’s d = 0.777), respectively, suggesting multisensory gains. More importantly, at 1Hz, the amplitude in the AV condition was significantly larger than the algebraic sum of those in the A and V conditions (t (23) = 3.028, p = 0.006, Cohen’s d = 0.618), indicating a super-additive audiovisual integration effect. While at 2Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = −0.623, p = 0.539, Cohen’s d = −0.127), indicating additive audiovisual integration.

Experiment 1b

To further test whether cortical entrainment can apply to stimuli with a different speed, Experiment 1b altered the frequencies of the gait cycle and the corresponding step cycle to 0.83 Hz and 1.67 Hz while adopting the same paradigm as Experiment 1a. Consistent with Experiment 1a, the frequency-domain analysis revealed significant cortical entrainment to the audiovisual stimuli at the new speeds. As shown in Fig. 2d, both the responses to V, A, and AV stimuli showed clear peaks at step-cycle frequency (1.67 Hz; V: t (23) = 3.473, p = .001; A: t (23) = 9.194, p < .001; AV: t (23) = 8.756, p < .001; FDR corrected) and its harmonics (3.33 Hz, ps < .001, FDR corrected, see Supplementary Information for additional analysis). In contrast, at gait-cycle frequency (0.83 Hz), only the response to AV stimuli showed significant peaks (V: t (23) = −1.125, p = .846; A: t (23) = −2.449, p = .989; AV: t (23) = 3.052, p = .003; FDR corrected).

At both 0.83 Hz (Fig. 2e) and 1.67 Hz (Fig. 2f), the amplitude in the AV condition was stronger or marginally stronger than that in the V condition (0.83 Hz: t (23) = 2.665, p = .014, Cohen’s d = 0.544; 1.67 Hz: t (23) = 6.380, p < .001, Cohen’s d = 1.302) and the A condition (0.83 Hz: t (23) = 3.625, p < .001, Cohen’s d = 0.740; 1.67 Hz: t (23) = 1.752, p = .093, Cohen’s d = 0.358), respectively, suggesting multisensory gains. More importantly, at 0.83 Hz, the amplitude in the AV condition was significantly larger than the sum of those in the A and V conditions (t (23) = 3.240, p = .004, Cohen’s d = 0.661), indicating a super-additive audiovisual integration effect. While at 1.67 Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = −0.735, p = .470, Cohen’s d = −0.150), indicating linear audiovisual integration.

In summary, results from Experiments 1a & 1b consistently showed that the cortical tracking of the audiovisual signals at different temporal scales exhibit distinct audiovisual integration modes, i.e., the super-additive effect at gait-cycle frequency and the additive effect at step-cycle frequency, indicating that the cortical entrainment effects at the two temporal scales might be driven by functionally different mechanisms.

Cortical tracking of higher-order rhythmic structure contributes to the AVI of BM

To further explore whether and how the cortical tracking of the rhythmic information contributes to the specialized audiovisual process of BM, both upright and inverted BM stimuli were adopted in Experiment 2. The task and the frequencies of visual stimuli in Experiment 2 were same as Experiment 1a. Specifically, participants were required to perform the change detection task when perceiving upright and inverted visual BM sequences (1 Hz for gait-cycle frequency and 2 Hz for step-cycle frequency) accompanied by frequency congruent (1 Hz) or incongruent (0.6 Hz and 1.4 Hz) footstep sounds. The audiovisual congruency effect, characterized by stronger neural responses in the audiovisual congruent condition compared with the incongruent condition, can be taken as an index of AVI (Fleming et al., 2020; Jones & Jarick, 2006; Maddox et al., 2015; Wuerger, Crocker-Buque, et al., 2012). A stronger congruency effect in the upright condition relative to the inverted condition characterizes the AVI process specific to BM information.

We calculated the audiovisual congruency effect for the upright (AVIupr) and the inverted (AVIinv) conditions, respectively. Then, we identified the clusters showing significantly different congruency effects between the upright and inverted conditions using a cluster-based permutation test over all electrodes (n = 1000, alpha = 0.05; see Methods). At 1 Hz, the congruency effect in the upright condition was significantly stronger than that in the inverted condition in a cluster at the right hemisphere (Fig. 3a, lower panel, p = 0.029; C2, CPz, CP2, CP4, CP6, Pz, P2, P4, P6), revealing a BM-specific AVI process. Then we averaged the amplitude of electrodes within the significant cluster and further performed a two-tailed paired t-test. The results showed that (Fig. 3b) the audiovisually congruent BM information enhanced the oscillatory amplitude relative to the incongruent ones only for upright BM stimuli (t (23) = 4.632, p < 0.001, Cohen’s d = 0.945) but not when visual BM was inverted (t (23) = 0.480, p = 0.635, Cohen’s d = 0.098). The congruency effect in the upright condition was significantly larger than that in the inverted condition (t (23) = 3.099, p = 0.005, Cohen’s d = 0.633).

Cortical tracking at gait-cycle rather than step-cycle frequency contributes to BM-specific AVI effect.

(a) & (d) Topographic maps for the congruency effect at 1Hz and 2 Hz for each condition: Upright, Inverted, and for the difference between these conditions (Upright-Inverted), respectively. The black dots indicate the electrodes showing a significant congruency effect (upper panels) or a significantly enhanced congruency effect in the upright condition relative to the inverted condition (lower panel). Then, the amplitude at the electrodes shown in the lower panel of (a) were averaged to quantify the cortical entrainment effect at 1 Hz (b) and 2 Hz (e). Error bars represent ±1 standard error of means. Individuals’ autistic traits correlated with the BM-specific AVI at 1 Hz (c) but not 2 Hz (f). Shaded regions indicate the 95% confidence intervals.

In contrast, at 2 Hz, no cluster showed a significantly different congruency effect between the upright and inverted conditions (Fig. 3d). We then conducted further analysis on the averaged amplitude of the electrodes marked in Fig. 3a (lower panel). A two-tailed paired t-test showed that both upright and inverted stimuli induced a significant congruency effect at 2 Hz (Fig. 3e; Upright: t (23) = 3.096, p = 0.005, Cohen’s d = 0.632; Inverted: t (23) = 2.672, p = 0.014, Cohen’s d = 0.545). The congruency effect between the upright and inverted conditions was not different (t (23) = 0.434, p = 0.668, Cohen’s d = 0.089), suggesting a comparable audiovisual congruency effect between two conditions at 2 Hz. Importantly, a three-way repeated-measures ANOVA with frequency (1 Hz vs. 2 Hz), orientation (upright vs. inverted), and audiovisual congruency (congruent vs. incongruent) as within-subject factors revealed a marginal significant three-way interaction (F (1,23) = 3.190, p = 0.087, = 0.122), further implying that the audiovisual integration processing of BM is different between 1 Hz and 2 Hz.

BM-specific cortical tracking correlates with autistic traits

Furthermore, we examined the link between individuals’ autistic traits and the neural responses underpinning the AVI of BM, measured by the difference of congruency effect between the upright and the inverted BM conditions, using Pearson correlation analysis. After removing one outlier (exceeded 3 SD), we observed an evident negative correlation between individuals’ AQ scores and their neural responses at 1 Hz (Fig. 3c, r = −0.493, p = 0.017) but not at 2 Hz (Fig. 3f, r = −0.158, p = .460). The lack of significant results at 2 Hz was not attributable to electrode selection bias based on the significant cluster at 1 Hz, as similar results were observed when we performed the analysis on electrodes within the clusters showing significant congruency effects at 2 Hz (see the control analysis in Supplementary Information for details).

Discussion

The current study investigated the neural implementation for the AVI of human BM information and its functional implications. We found that, even under a motion-irrelevant color detection task, neural oscillations of observers entrained to temporally corresponding audiovisual BM signals at the frequencies of two rhythmic structures, i.e., the higher-order structure of gait cycle at a larger integration window and the basic-level structure of step cycle at a smaller integration window. Moreover, the strength of these cortical entrainment effects was enhanced under the audiovisual condition than in the visual-only or auditory-only condition, indicating multisensory gains in the cortical tracking of BM information (Experiments 1a & 1b).

Crucially, although the entrainment processes at both gait-cycle frequency and step-cycle frequency gain benefits from multisensory correspondence, the mechanisms underlying these two processes appear to be different. At step-cycle frequency, the cortical entrainment effect in the AV condition equals the additive sum of the unisensory conditions. Such linear integration might result from concurrent, independent processing of unisensory inputs without additional interaction of them (Stein et al., 2009). In contrast, at gait-cycle frequency, the congruent audiovisual signals led to a super-additive multisensory enhancement over the linear combination of auditory and visual conditions (AV > A+V), despite that there was no evident cortical tracking effect in the visual condition, different from previous findings obtained with a motion-relevant change detection task (Shen, Lu, Yuan, et al., 2023). This multisensory enhancement may bring about decreased thresholds of detection and identification (Stanford et al., 2005), allowing us to achieve a more clear and stable perception of the external environment and detect weak stimulus changes in time and respond adaptively.

Furthermore, results from Experiment 2 demonstrated that the cortical entrainment to gait-cycle rather than step-cycle is specific to the AVI of BM. In particular, the AVI effect at step-cycle frequency was significant for both upright and inverted BM signals and comparable between the two conditions, while the AVI effect at gait-cycle frequency was only significant in the upright condition and was greater than that in the inverted condition. These findings suggest that the cortical entrainment at step-cycle frequency reflects the integration of basic motion signals and corresponding sounds, while the cortical entrainment at gait-cycle frequency reflects the AVI of higher-level BM information. Together, these results reveal that the neural tracking of different levels of kinematic structures plays distinct roles in the AVI of BM, which may result from the interplay of stimulus-driven and domain-specific mechanisms.

Besides the temporal dynamics of neural activity revealed by the cortical entrainment process, we found that the BM-specific AVI effect was associated with enhanced cortical tracking of gait cycles in the right temporoparietal electrodes. This finding likely relates to neural activity in the right posterior superior temporal sulcus (pSTS), a region responding to both auditory and visual BM information and being causally involved in BM perception (Bidet-Caulet et al., 2005; Grossman et al., 2005; Wang et al., 2022). While previous fMRI studies have observed STS activation when processing spatial or semantic correspondence between audiovisual BM (Meyer et al., 2011; Wuerger, Parkes, et al., 2012), whether this region also engages in the audiovisual processing of BM signals based on temporal correspondence remains unknown. The current study provides preliminary evidence for such a possibility, inviting future research to localize the exact source of the multisensory integration processes based on imaging data with high temporal resolution and spatial resolution, such as MEG.

In a broad sense, the current study deepens our understanding of the neural processing of audiovisual signals in natural stimuli with complex temporal structures. Cortical entrainment can track simple rhythmic stimuli like tone sequences or luminance-varying patches (C. Keitel et al., 2017; Yuan et al., 2021) as well as complex rhythmic structures in speech (Brookshire et al., 2017; Ding et al., 2016; Keitel et al., 2018) and BM (Shen, Lu, Yuan, et al., 2023). Beyond unisensory processing, cortical entrainment also plays a role in the multisensory processing of simple or discrete rhythmic signals generated by physical stimulation (Bauer et al., 2021; Miller et al., 2013; Nozaradan et al., 2012b; Simon & Wallace, 2017). These findings may partially explain the AVI effect at 2 Hz for BM and non-BM stimuli in the current study. However, we found that the cortical tracking of the perceived higher-order rhythmic structure based on spatiotemporal integration of meaningful BM information (i.e., the gait cycle of upright walkers rather than inverted walkers) is selectively engaged by the AVI of BM, suggesting that the multisensory processing of natural continuous stimuli may involve unique mechanisms besides the purely stimulus-driven AVI process. Similar to BM, other natural rhythmic stimuli, like auditory speech, also convey hierarchical structures that can entrain neural oscillations at different temporal scales (Ding et al., 2016; Keitel et al., 2018). Previous studies have observed the AVI of speech at theta band (4-6 Hz), a temporal scale that corresponds to the rate of syllables (Crosse et al., 2015), and that the asynchrony detection of prosodic fluctuation in audiovisual speech linked with delta oscillations (∼1-3 Hz) (Biau et al., 2022). These findings raise a possibility that the AVI of speech also occurs at multiple temporal scales and that the multi-scale entrainment effects play different roles in speech perception. Further investigation into these issues and comparing the results with BM studies will help complete the picture of how the human brain integrates complex, rhythmic information sampled from different sensory modalities to orchestrate perception in a natural scenario.

Last but not least, our study demonstrated that the selective cortical tracking of higher-level rhythmic structure in audiovisually congruent BM signals negatively correlated with individual autistic traits. This finding highlights the functional significance of cortical tracking and integration of audiovisual BM signals in social cognition. It also offers the first evidence that differences in audiovisual BM processing are already present in nonclinical individuals and associated with their autistic traits, beyond previous evidence for atypical audiovisual BM processing in ASD populations (Falck-Ytter et al., 2013, 2018; Klin et al., 2009), lending support to the continuum view of ASD (Baron-Cohen et al., 2001). Meanwhile, given that impaired audiovisual BM processing at the early stage may influence social development and result in cascading consequences for lifetime impairments in social interaction (Falck-Ytter et al., 2018; Klin et al., 2005), it is worth exploring neural entrainment to the temporal correspondence of audiovisual BM signals in children with different autistic levels, which may help reveal whether deficits in such ability could serve as an early neural hallmark for ASD.

Materials and Methods

Participants

Seventy-two participants (mean age ± SD = 22.4 ± 2.6 years, 35 females) took part in the study, 24 for each of Experiment 1a, Experiment 1b, and Experiment 2. All of them had normal or corrected-to-normal vision and reported no history of neurological, psychiatric, or hearing disorders. They were naïve to the purpose of the study and gave informed consent according to procedures and protocols approved by the institutional review board of the Institute of Psychology, Chinese Academy of Sciences.

Stimuli

Visual stimuli

The visual stimuli (Fig. 1a, left panel) consisted of 13 point-light dots attached to the head and major joints of a human walker (Vanrie & Verfaillie, 2004). This point-light walker looked like walking on a treadmill and did not translate on the screen. It conveys rhythmic structures specified by recurrent forward motions of bilateral limbs (Fig. 1a, right panel). Each step, regardless of left or right foot, occurs recurrently to form a step cycle. The antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, a full gait cycle took 1 second and was repeated 6 times to form a 6-second walking sequence. That is, the gait-cycle frequency is 1 Hz and the step-cycle frequency is 2 Hz. In Experiment 1b, the gait-cycle frequency was 0.83 Hz and the step-cycle frequency was 1.67 Hz. The gait cycle was repeated 6 times to form a 7.2-second walking sequence. The stimuli in Experiment 2 were the same as that in Experiment 1a. Meanwhile, the point-light BM was mirror-flipped vertically to generate inverted BM (Fig. 1a, left panel), which preserves the temporal structure of the stimuli but distorts its distinctive kinematic features, such as movement that is compatible with the effect of gravity (Shen, Lu, Yuan, et al., 2023; Troje & Westhoff, 2006; Wang et al., 2022).

Auditory stimuli

Auditory stimuli were continuous footstep sounds (6 s) with a sampling rate of 44,100 Hz. As shown in Fig. 1b, in Experiments 1a & 2, the gait-cycle frequency of congruent sounds was 1 Hz, which consisted of two steps or two impulses generated by each foot striking the ground within one gait cycle. The incongruent sounds included a faster (1.4 Hz) and a slower (0.60 Hz) sound. Both congruent and incongruent sounds were generated by manipulating the temporal interval between two successive impulses based on the same auditory stimuli. In Experiment 1b, the gait-cycle frequency of sound was 0.83 Hz.

Stimuli presentation

The visual stimuli were rendered white against a grey background and displayed on a CRT (cathode ray tube) monitor. Participants sat 60 cm from the computer screen (1280×1024 at 60 Hz; High: 37.5 cm; Width: 30 cm), with their heads held stationary on a chinrest. The auditory stimuli were presented binaurally over insert earphones. All stimuli were generated and presented using MATLAB together with the Psychophysics Toolbox (Brainard, 1997; Pelli, 1997).

Procedure and task

Experiment 1a

The experiment was conducted in an acoustically dampened and electromagnetically shielded chamber. Participants completed the task under three conditions (Visual: V; Auditory: A; Audiovisual: AV) with the same procedure (Fig. 1c) except for the stimuli. In the V condition, each trial began with a white fixation cross (0.42° × 0.42°) displayed at the center of a gray background for a random duration (0.8 s to 1 s). Subsequently, a 6-s point-light walker (3.05°×5.47°) walked toward the left or right at a constant walking cycle frequency (1 Hz). To maintain observers’ attention, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed (the RGB values changed from [255 255 255] to [207 207 207]) one or two times throughout the trial. Each change lasted 0.5 s. Observers were required to report the number of changes (0, 1, or 2) via keypresses as accurately as possible after the point-light display was replaced by a red fixation. The next trial started 2–3 s after the response. In the A condition, the 6 s-stimuli were replaced by a visually static BM figure accompanied by continuous footstep sounds. The frequency of footstep sounds was congruent with the frequency of visual BM in the V condition. In the AV condition, the stimuli were temporally congruent visual BM sequences (as in the V condition) and footstep sounds (as in the A condition). Three conditions were conducted in separate blocks. V condition was performed in the middle of A and AV conditions. The order of A and AV conditions was counterbalanced across participants. Each participant completed 40 experimental trials without changes and 10-15 catch trials in each condition, resulting in a total of 150-165 trials. In each condition, participants completed a practice session with 3 trials to get familiar with the task before the formal EEG experiment.

Experiment 1b

The procedure of Experiment 1b was the same as that for Experiment 1a but with two exceptions. First, to test if the cortical entrainment effect can apply to stimuli with a different speed, we altered the frequencies of gait and step cycles to 0.83 Hz and 1.67 Hz. Second, we presented the 3 conditions (V, A, and AV) in a completely random order to eliminate the influence of presentation order. To minimize the potential influence of condition switch, we increased the trial number in the practice session from 3 to 14 for each condition.

Experiment 2

The procedure in Experiment 2 was similar to the AV condition in Experiment 1a, except that the visually displayed BM was accompanied by frequency congruent (1 Hz) or incongruent (0.6 or 1.4 Hz) footstep sounds. Each participant completed a total of 76 experiment trials, consisting of 36 congruent-trials, 20 incongruent-trials with faster sounds (1.4 Hz), and 20 incongruent-trials with slower sounds (0.6 Hz). These trials were assigned to 3 blocks based on the frequency of the footstep sounds, with the order of the three frequencies balanced across participants. Besides, an inverted BM was used as a control to investigate whether there is a specialized mechanism tuned to the AVI of life motion signals. The order of upright and inverted conditions was balanced across participants. Meanwhile, we measured the participants’ autistic traits by using the Autism-Spectrum Quotient, or AQ questionnaire (Baron-Cohen et al., 2001). Higher AQ scores indicate a higher level of autistic traits.

EEG recording and analysis

EEG was recorded at 1000 Hz using a SynAmps2 NeuroScan amplifier System with 64 electrodes placed on the scalp according to the international 10-20 system. Horizontal and vertical eye movements were measured via four additional electrodes placed on the outer canthus of each eye and the inferior and superior areas of the left orbit. Impedances were kept below 5 kΩ for all electrodes.

Preprocessing

The catch trials were excluded from EEG analysis. All preprocessing and further analyses were performed using the FieldTrip toolbox (Maris & Oostenveld, 2007) in the MATLAB environment. EEG recordings were pass-filtered between 0.1 and 30 Hz, and down-sampled to 100 Hz. Then the continuous EEG data were cut into epochs ranging from −1s to 6 gait cycles (7.2 s in Experiment 1b and 6 s in other experiments) time-locked to the onset of the visual point-light stimuli. The epochs were visually inspected, and trials contaminated with excessive noise were excluded from the analysis. After the trial rejection, eye and cardiac artifacts were removed via independent component analysis based on the Runica algorithm (Bell & Sejnowski, 1995; Jung et al., 2000; Makeig, 2002). Then the cleaned data were re-referenced to the average mastoids (M1 and M2). To minimize the influence of stimulus-onset evoked activity on EEG spectral decomposition, the EEG recording before the onset of the stimulus and the first cycle (1 s in Experiments 1a & 2; 1.2 s in Experiment 1b) of each trial was excluded (Nozaradan et al., 2012a). After that, the EEG epochs were averaged across trials for each participant and condition.

Frequency-Domain analysis and statistics

A Fast Fourier Transform (FFT) with zero padding (1200) was used to convert the averaged EEG signals from the temporal domain to the spectral domain, resulting in a frequency resolution of 0.083 Hz, i.e., 1/12 Hz, which is sufficient for observing neural responses around the frequency of the rhythmic BM structures in all experiments. When performing FFT, a Hanning window was adopted to minimize spectral leakage. Then, to remove the 1/f trend of the response amplitude spectrum and identify spectral peaks, the response amplitude at each frequency was normalized by subtracting the average amplitude measured at the neighboring frequency bins (two bins on each side) (Nozaradan et al., 2012a). We calculated the normalized amplitude separately for each electrode (except for electrooculogram electrodes, CB1, and CB2), participant, and condition.

In Experiment 1, the normalized amplitude in all electrodes was averaged and a right-tailed one-sample t-test against zero was performed on the grand average amplitude to test whether the neural response in each frequency bin showed a significant entrainment effect or spectral peak. This test was applied to all frequency bins below 5.33 Hz and multiple comparisons were controlled by false discovery rate (FDR) correction at p < 0.05 (Benjamini & Hochberg, 1995). In Experiment 2, to further identify the BM-specific AVI process, the audiovisual congruency effect was compared between the upright and inverted conditions using a cluster-based permutation test over all electrodes (1000 iterations, requiring a cluster size of at least 2 significant neighbors, a two-sided t-test at p < 0.05 on the clustered data) (Maris & Oostenveld, 2007). This allowed us to identify the spatial distribution of the BM-specific congruency effect.

eLife Assessment

Wang et al. presented visual (dot) motion and/or the sound of a walking person and found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm, with tentative demonstration that the gait rhythm is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition). The findings will be of wide interest to those examining biological motion perception and oscillatory processes more broadly, with the potential to be important. However, at present, due to some analysis concerns - most notably, evidence of double-dipping for one of the core findings - the evidence is incomplete. Furthermore, some of the theoretical interpretations concerning entrainment must remain speculative when the authors cannot dissociate evoked responses from entrained oscillatory effects.

Reviewer #1 (Public Review):

Summary:

Shen et al. conducted three experiments to study the cortical tracking of the natural rhythms involved in biological motion (BM), and whether these involve audiovisual integration (AVI). They presented participants with visual (dot) motion and/or the sound of a walking person. They found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm. The gait rhythm specifically is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition, Experiments 1a/b), which is independent of the specific step frequency (Experiment 1b). Furthermore, audiovisual integration during tracking of gait was specific to BM, as it was absent (that is, the audiovisual congruency effect) when the walking dot motion was vertically inverted (Experiment 2). Finally, the study shows that an individual's autistic traits are negatively correlated with the BM-AVI congruency effect.

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

Weaknesses:

I only have one potential worry. The analysis for gait tracking (1 Hz) in Experiment 2 (Figures 3a/b) starts by computing a congruency effect (A/V stimulation congruent (same frequency) versus A/V incongruent (V at 1 Hz, A at either 0.6 or 1.4 Hz), separately for the Upright and Inverted conditions. Then, this congruency effect is contrasted between Upright and Inverted, in essence computing an interaction score (Congruent/Incongruent X Upright/Inverted). Then, the channels in which this interaction score is significant (by cluster-based permutation test; Figure 3a) are subselected for further analysis. This further analysis is shown in Figure 3b and described in lines 195-202. Critically, the further analysis exactly mirrors the selection criteria, i.e. it is aimed at testing the effect of Congruent/Incongruent and Upright/Inverted. This is colloquially known as "double dipping", the same contrast is used for selection (of channels, in this case) as for later statistical testing. This should be avoided, since in this case even random noise might result in a significant effect. To strengthen the evidence, either the authors could use a selection contrast that is orthogonal to the subsequent statistical test, or they could skip either the preselection step or the subsequent test. (It could be argued that the test in Figure 3b and related text is not needed to make the point - that same point is already made by the cluster-based permutation test.)

Related to the above: the test for the three-way interaction (lines 211-216) is reported as "marginally significant", with a p-value of 0.087. This is not very strong evidence.

Reviewer #2 (Public Review):

Summary:

The authors evaluate spectral changes in electroencephalography (EEG) data as a function of the congruency of audio and visual information associated with biological motion (BM) or non-biological motion. The results show supra-additive power gains in the neural response to gait dynamics, with trials in which audio and visual information were presented simultaneously producing higher average amplitude than the combined average power for auditory and visual conditions alone. Further analyses suggest that such supra-additivity is specific to BM and emerges from temporoparietal areas. The authors also find that the BM-specific supra-additivity is negatively correlated with autism traits.

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

Weaknesses:

The manuscript interprets the neural findings using mechanistic and cognitive claims that are not justified by the presented analyses and results.

First, entrainment and cortical tracking are both invoked in this manuscript, sometimes interchangeably so, but it is becoming the standard of the field to recognize their separate evidential requirements. Namely, step and gate cycles are striking perceptual or cognitive events that are expected to produce event-related potentials (ERPs). The regular presentation of these events in the paradigm will naturally evoke a series of ERPs that leave a trace in the power spectrum at stimulation rates even if no oscillations are at play. Thus, the findings should not be interpreted from an entrainment framework except if it is contextualized as speculation, or if additional analyses or experiments are carried out to support the assumption that oscillations are present. Even if oscillations are shown to be present, it is then a further question whether the oscillations are causally relevant toward the integration of biological motion and for the orchestration of cognitive processes.

Second, if only a cortical tracking account is adopted, it is not clear why the demonstration of supra-additivity in spectral amplitude is cognitively or behaviorally relevant. Namely, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated with the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion, and it is possible that such neural correlates relate to behaviorally relevant neural mechanisms, but based on the current task and associated analyses this has not been shown.

Reviewer #3 (Public Review):

Summary:

The study demonstrates differential patterns of entrainment to biological motion (BM). At a basic, sensory level, the authors demonstrate entrainment to faster rhythms that make up BM (step-cycle) which seems to be separate from its audio aspects and its visual aspects (though to a much lesser degree). Ultimately this temporal scale seems to reside in a manner that does not indicate much multi-modal integration. At a higher-order, emergent rhythms in motion that are biologically relevant (gait-cycle) seem to be the result of multisensory integration. The work sheds light on the perceptual processes that are engaged in perceiving BM as well as the role of multisensory integration in these processes. Moreover, the work also outlines interesting links between shorter and longer integration windows along the sensory and multisensory processing stages.

In a series of experiments, the authors sought to investigate the role of multisensory integration in the processing of biological motion (BM). Specifically, they study neural entrainment in BM light-point walkers. Visual-only, auditory-only, and audio-visual (AV) displays were compared under different conditions.

Experiments 1a and b mainly characterized entrainment to these stimuli. Here, entrainment to step cycle (at different scales for 1a and 1b) was found to entrain in the presence of the auditory rhythm and to a certain degree also for the visual stimulus (though barely beyond the noise floor in 1b). The AV condition for this temporal scale seemed to follow an additive rule whereby the combined stimulation resulted in entrainment more or less equal to the sum of the unimodal effects. At the slower, gait cycle a slightly different pattern emerges whereby neither unimodal stimulation conditions result in entrainment however the AV condition does.

This finding was further explored in Experiment 2 where two extra manipulations were added. Point-light walkers could generally be either congruently paired with AV or incongruently. In addition, the visual BM stimulus was matched with a control consisting of an inverted BM and thus non-BM movement. This study enabled further discerning among the step- and gait-cycle findings seeing that the pattern that emerged suggested that step-cycle entrainment was consistent with a low-level process that is not selective to BM whilst gait-cycle entrainment was only found for BM. This generally replicated the findings in Experiment 1 and extended them further suggesting that entrainment seen for uni- and multisensory step cycles is reflects a different process than that captured in the gait-cycle multi-modal entrainment. The selective BM finding seemed to demonstrate a link to autistic traits within a sample of 24 participants informing a hypothesis that sensitivity to biological motion might be related to social cognition.

Strengths:

The main strengths of the paper relate to the conceptualization of BM and the way it is operationalized in the experimental design and analyses. The use of entrainment, and the tracking of different, nested aspects of BM result in seemingly clean data that demonstrate the basic pattern. The first experiments essentially provide the basic utility of the methodological innovation and the second experiment further hones in on the relevant interpretation of the findings by the inclusion of better control stimuli sets.

Another strength of the work is that it includes at a conceptual level two replications.

Weaknesses:

The statistical analysis is misleading and inadequate at times. The inclusion of the autism trait is not foreshadowed and adequately motivated and is likely underpowered. Finally, a broader discussion over other nested frequencies that might reside in the point-light walker stimuli would also be important to fully interpret the different peaks in the spectra.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #1 (Public Review):

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

We are grateful to the reviewer’s thoughtful comments.

Weaknesses:

I only have one potential worry. The analysis for gait tracking (1 Hz) in Experiment 2 (Figures 3a/b) starts by computing a congruency effect (A/V stimulation congruent (same frequency) versus A/V incongruent (V at 1 Hz, A at either 0.6 or 1.4 Hz), separately for the Upright and Inverted conditions. Then, this congruency effect is contrasted between Upright and Inverted, in essence computing an interaction score (Congruent/Incongruent X Upright/Inverted). Then, the channels in which this interaction score is significant (by cluster-based permutation test; Figure 3a) are subselected for further analysis. This further analysis is shown in Figure 3b and described in lines 195-202. Critically, the further analysis exactly mirrors the selection criteria, i.e. it is aimed at testing the effect of Congruent/Incongruent and Upright/Inverted. This is colloquially known as "double dipping", the same contrast is used for selection (of channels, in this case) as for later statistical testing. This should be avoided, since in this case even random noise might result in a significant effect. To strengthen the evidence, either the authors could use a selection contrast that is orthogonal to the subsequent statistical test, or they could skip either the preselection step or the subsequent test. (It could be argued that the test in Figure 3b and related text is not needed to make the point - that same point is already made by the cluster-based permutation test.)

Thanks for the helpful suggestions. In Experiment 2, to investigate whether the multisensory integration effect was specialized for biological motion perception, we contrasted the congruency effect between the upright and inverted conditions to search for clusters showing a significant interaction effect. We performed further analyses based on neural responses from this cluster to examine whether the congruency effect was significant in the upright and the inverted conditions, respectively, following the logic of post hoc comparisons after identifying an interaction effect. However, we agree with the reviewer that comparing the congruency effects between the upright and inverted conditions again based on data from this cluster was redundant and resulted in doubledipping. Therefore, we have removed this comparison from the main text and optimized the way to present our results in the revised Fig. 3).

Related to the above: the test for the three-way interaction (lines 211-216) is reported as "marginally significant", with a p-value of 0.087. This is not very strong evidence.

As shown in Fig.3b & e, the magnitude of amplitude differs between the gaitcycle frequency (mean = 0.008, SD = 0.038) and the step-cycle frequency (mean = 0.052; SD =0.056), which might influence the statistical results of the interaction effect. To reduce such influence, we converted the amplitude data at each frequency condition into Z-scores, separately. The repeated-measures ANOVA analysis on these normalized amplitude data revealed a significant three-way interaction (F (1,23) = 7.501, p = 0.012, ƞp2 = 0.246). We have updated the results in the revised manuscript (lines 218-225).

Reviewer #1 (Recommendations For The Authors):

- Which variable caused one data point to be classified as outlier? (line 221).

The outlier is a participant whose audiovisual congruency effect (Upright – Inverted) in neural responses at the frequency of interest exceeds 3 SD from the group mean. It is marked by a red diamond in Author response 2. Before removing the data, the correlation between the AQ score and the congruency effect is r = -0.396, p = 0.055. For comparison, the results after removing the outlier are shown in Fig. 3c of the revised manuscript. We have added more information about the variable causing the outlier in the revised manuscript (lines 231-232).

The correlation between AQ score and congruency effect

- The authors cite Maris & Oostenveld (2007) in line 415 as the main reference for the FieldTrip toolbox, but the correct reference here is different, see https://www.fieldtriptoolbox.org/faq/how_should_i_refer_to_fieldtrip_in_my_p ublication/

Thank you for pointing out this issue. Citation corrected.

- The authors could consider giving some more background on the additive vs superadditive distinction in the Introduction, which may increase the impact; as it stands the reader might not know why this is particularly interesting. Summarize some of the takeaways of the Stevenson et al. (2014) review in this respect.

Thanks for the suggestion and we have added the following relevant information in the Introduction (lines 80-90):

“Moreover, we adopted an additive model to classify multisensory integration based on the AV vs A+V comparison. This model assumes independence between inputs from each sensory modality and distinguishes among sub-additive (AV < A+V), additive (AV = A+V), and super-additive (AV > A+V) response modes (see a review by Stevenson et al., 2014). The additive mode represents a linear combination between two modalities. In contrast, the super-additive and subadditive modes indicate non-linear interaction processing, either with potentiated neural activation to facilitate the perception or detection of nearthreshold signals (super-additive) or a deactivation mechanism to minimize the processing of redundant information cross-modally (sub-additive) (Laurienti et al., 2005; Metzger et al., 2020; Stanford et al., 2005; Wright et al., 2003).”

Reviewer #2 (Public Review):

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

We thank the reviewer for the valuable feedback.

Weaknesses:

The manuscript interprets the neural findings using mechanistic and cognitive claims that are not justified by the presented analyses and results.

First, entrainment and cortical tracking are both invoked in this manuscript, sometimes interchangeably so, but it is becoming the standard of the field to recognize their separate evidential requirements. Namely, step and gate cycles are striking perceptual or cognitive events that are expected to produce event-related potentials (ERPs). The regular presentation of these events in the paradigm will naturally evoke a series of ERPs that leave a trace in the power spectrum at stimulation rates even if no oscillations are at play. Thus, the findings should not be interpreted from an entrainment framework except if it is contextualized as speculation, or if additional analyses or experiments are carried out to support the assumption that oscillations are present. Even if oscillations are shown to be present, it is then a further question whether the oscillations are causally relevant toward the integration of biological motion and for the orchestration of cognitive processes.

Second, if only a cortical tracking account is adopted, it is not clear why the demonstration of supra-additivity in spectral amplitude is cognitively or behaviorally relevant. Namely, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated with the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion, and it is possible that such neural correlates relate to behaviorally relevant neural mechanisms, but based on the current task and associated analyses this has not been shown.

Thanks for raising the important concerns regarding the interpretation of our results within the entrainment or the cortical tracking frame. A strict neural entrainment account emphasizes the alignment of endogenous neural oscillations with external rhythms, rather than a mere regular repetition of stimulus-evoked responses. However, it is challenging to fully dissociate these components, given that rhythmic stimulation can shape intrinsic neural oscillations, resulting in an intricate interplay between endogenous neural oscillations and stimulus-evoked responses (Duecker et al., 2024; Herrmann et al., 2016; Hosseinian et al., 2021). Therefore, some research, including the current study, use the term “entrainment” to refer to the alignment of brain activity to rhythmic stimulation in a broader context, without isolating the intrinsic oscillations and evoked responses (e.g., Ding et al., 2016; Nozaradan et al., 2012; Obleser & Kayser, 2019). Nevertheless, we agree with the reviewer that since the current results did not examine or provide direct evidence for endogenous oscillations, it is better to contextualize the oscillation view as speculations. Hence, we have replaced most of the expressions about “entrainment” with a more general term “tracking” in the revised manuscript (as well as in the title of the manuscript). We only briefly mentioned the entrainment account in the Discussion to facilitate comparison with the literature (lines 307-312).

Regarding the relevance between neural findings and cognition or behavioral performance, the first supporting evidence comes from the inversion effect in Experiment 2. For the neural responses at gait-cycle frequency, we observed a significantly enhanced audiovisual congruency effect in the upright condition compared with the inverted condition. Inversion disrupts the distinctive kinematic features of biological motion (e.g., gravity-compatible ballistic movements) and significantly impairs biological motion processing, but it does not change the basic visual properties of the stimuli, including the rhythmic signals generated by low-level motion cues. Therefore, the inversion effect has long been regarded as an indicator of the specificity of biological motion processing in numerous behavioral and neuroimaging studies (Bardi et al., 2014; Grossman & Blake, 2001; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Vallortigara & Regolin, 2006; Wang et al., 2014; Wang & Jiang, 2012; Wang et al., 2022). Here, our finding of the cortical tracking of higher-order rhythmic structures (gait cycles) present in the upright but not in the inverted condition suggests that this cortical tracking effect can not be explained by ERPs evoked by regular onsets of rhythmic events. Rather, it is closely linked with the specialized cognitive processing of biological motion. Furthermore, we found that the BM-specific cortical tracking effect at gait-cycle frequency (rather than the non-selective tracking effect at step-cycle frequency) correlates with observers’ autistic traits, indicating its functional relevance to social cognition. These findings convergingly suggest that the cortical tracking effect that we currently observed engages cognitively relevant neural mechanisms. In addition, our recent behavioral study showed that listening to frequency-congruent footstep sounds, compared with incongruent sounds, enhanced the visual search for human walkers but not for non-biological motion stimuli containing the same rhythmic signals (Shen, Lu, Wang, et al., 2023). These results suggest that audiovisual correspondence specifically enhances the perceptual and attentional processing of biological motion. Future research could examine whether the cortical tracking of rhythmic structures plays a functional role in this process, which may shed more light on the behavioral relevance of the cortical tracking effect to biological motion perception. We have incorporated the above information into the Discussion (lines 268-293).

Reviewer #2 (Recommendations For The Authors):

In Figure 1c, it could be helpful to add the word "static" in the illustration for the auditory condition so that readers understand without reading the subtext that it is a static image without biological motion.

Suggestion taken.

In the Discussion, I believe it is important to justify an oscillation and entrainment account, or if it cannot be justified based on the current results and analyses (which is my opinion), it could be helpful to explicitly frame it as speculation.

We agree with the reviewer. For more clarification, please refer to our response to the public review.

L335, I did not understand this sentence - a reformulation would be helpful.

The point-light stimuli were created by capturing the motion of a walking actor (Vanrie & Verfaillie, 2004). The global motion of the walking sequences was eliminated so that the point-light walker looks like walking on a treadmill without translational motion. We have reformulated the sentence as follows: “The point-light walker was presented at the center of the screen without translational motion.”

The results in Figure 2a and 2d are derived by performing a t-test between the amplitude at the frequency of gait and step cycles and zero. Comparison against amplitude of zero is too liberal; the possibility for a Type-I error is inflated because even EEG data with only noise will not have amplitudes of zero at all frequencies. A better baseline (H0) is either the 1/frequency trend in the power spectrum derived using methods like FOOOF (https://fooof-tools.github.io/fooof/) or by performing non-parametric shuffling based methods (https://doi.org/10.1016/j.jneumeth.2007.03.024).

In our data analysis, instead of performing the t-test between raw amplitude with zero, we compared the normalized amplitude at each frequency bin (by subtracting the average amplitude measured at the neighboring frequency bins from the original amplitude data) against zero. Such analysis is equal to contrasting the raw amplitude to its neighboring frequency bins, allowing us to test whether the neural response in each frequency bin showed a significant enhancement compared with its neighbors. The multiple comparisons on each frequency bin were controlled by false discovery rate (FDR) correction, reducing the Type-I error. Such analysis procedures help reduce (though not totally remove) the influence of the 1/f trend and have been widely used in this field (Cirelli et al., 2016; Henry & Obleser, 2012; Lenc et al., 2018; Nozaradan et al., 2012; Peter et al., 2023).

To further verify our findings, we adopted the reviewer’s suggestion and created a baseline by performing a non-parametric shuffling-based analysis. More specifically, to establish the statistical significance of amplitude peaks, we carried out a surrogate analysis on each condition. For each participant, a single control surrogate dataset was derived from their actual dataset by jittering the onset of each step-cycle relative to the actual original onset by a randomly selected integer value ranging between − 490–490 ms. This procedure removed the consistent relationship between the EEG signal and the stimuli while preserving each epoch’s general timing within the exposure period. Then, epochs were extracted based on surrogate stimuli onset, and amplitude was computed across frequencies through FFT under a null model of non-entrainment (Moreau et al., 2022). This entire procedure was performed 100 times, producing a surrogate amplitude distribution of 100 group-averaged values for each condition. If the observed amplitude values at the frequency of interest exceeded the value corresponding to the 95th percentile of the surrogate distribution (p < .05) within a given condition (e.g., AV), the amplitude peak was considered significant (Batterink, 2020). As shown in Author response image 2, the statistical results from these analyses are similar to those reported in the manuscript, confirming the significant amplitude peaks at the frequencies of interest.

Non-parametric analysis for spectral peak.

The dotted lines represent the random data based on shuffling analysis. The solid lines represent the observed data in measured EEG signals. All conditions induced significant peaks at step-cycle frequency and its harmonic, while only the AV condition induced a significant peak at gait-cycle frequency.

Reviewer #3 (Public Review):

Strengths:

The main strengths of the paper relate to the conceptualization of BM and the way it is operationalized in the experimental design and analyses. The use of entrainment, and the tracking of different, nested aspects of BM result in seemingly clean data that demonstrate the basic pattern. The first experiments essentially provide the basic utility of the methodological innovation and the second experiment further hones in on the relevant interpretation of the findings by the inclusion of better control stimuli sets.

Another strength of the work is that it includes at a conceptual level two replications.

We appreciate the reviewer for the comprehensive review and positive comments.

Weaknesses:

The statistical analysis is misleading and inadequate at times. The inclusion of the autism trait is not foreshadowed and adequately motivated and is likely underpowered. Finally, a broader discussion over other nested frequencies that might reside in the point-light walker stimuli would also be important to fully interpret the different peaks in the spectra.

(1) Regarding the nested frequency peaks in the spectra, we did observe multiple significant amplitude peaks at 1f (1/0.83 Hz), 2f (2/1.67 Hz), and 4f (4/3.33 Hz) relative to the gait-cycle frequency (Fig. 2 a&d). To further test the functional roles of the neural activity at different frequencies, we analyzed the audiovisual integration modes at each frequency. Note that we collapsed the data from Experiments 1a & 1b in the analysis as they yielded similar results. Overall, results show a similar additive audiovisual integration mode at 2f and 4f and a super-additive integration mode only at 1f (Figure S1), suggesting that the cortical tracking effects at 2f and 4f may be functionally linked but independent of that at 1f. We have reported the detailed results in the Supplementary Information.

(2) For the reviewer’s other concerns about statistical analysis and autism traits, please refer to our responses below to the Recommendations for the authors.

Reviewer #3 (Recommendations For The Authors):

The description of the analyses performed for experiment 2 comes across as double dipping. Congruency effects for BM and non-BM motion (inverted) were compared using cluster-based statistics. Then identified clusters informed an averaging of signals which then were subjected to a paired comparison. At this point, it is no surprise that these paired comparisons are highly significant seeing that the channels were selected based on a cluster analysis of the same exact contrast. This approach should be avoided.

In the analysis of the repeated measures ANOVA reporting a trend as marginally significant is misleading. Reporting the statistical results whilst indicating that those do not reach significance is the appropriate way to communicate this finding. Other statistics can be used in order to provide the likelihood of those findings supporting H1 or H0 if the authors would like to state something more precise (Bayesian).

Thanks for the comments. We have addressed these two points in our response to the public review of Reviewer #1.

The authors perform a correlation along "autistic trait" scores in an individual differences approach. Individual differences are typically investigated in larger samples (>n=40). In addition, the range of AQ scores seems limited to mostly average or lower-than-average AQs (barring a couple). These points make the conclusions on the possible role of BM in the autistic phenotype very tentative. I would recommend acknowledging this.

An alternative analysis approach that might better suit the smaller sample size is a comparison between high and low AQ participants, defined based on a median split.

Many thanks for the suggestion. We agree with the reviewer that the sample size (n = 24) in the current study is not large for exploring the correlation between BM and autistic traits. The narrow range of AQ scores was due to the fact that all participants were non-clinical populations and we did not pre-select participants by AQ scores. To further confirm our findings, we adopted your suggestion to compare the BM-specific cortical tracking effect (i.e., audiovisual congruency effect (Upright - Inverted)) between high and low AQ participants split by the median AQ score (20) of this sample. Similar to correlation analysis, one outlier, whose audiovisual congruency effect (Upright – Inverted) in neural responses at 1 Hz exceeds 3 SD from the group mean, was removed from the following analysis. As shown in Figure S3, at 1 Hz, participants with low AQ showed a greater cortical tracking effect compared with high AQ participants (t (21) = 2.127, p = 0.045). At 2 Hz, low and high AQ participants showed comparable neural responses (t (22) = 0.946, p = 0.354). These results are in line with the correlation analysis, providing further support to the functional relevance between social cognition and cortical tracking of biological motion as well as its dissociation at the two temporal scales. We have added these results to the main text (lines 238-244) and the supplementary information.

Writing

The narrative could be better unfolded and studies better motivated. The transition from basic science research on BM to possibly delineating a mechanistic understanding of autism was a surprise at the end of the intro. Once the authors consider the suggestions and comments above it would be good to have this detail and motivation more obviously foreshadowed in the text.

Thanks for the great suggestion and we have provided an introduction about how audiovisual BM processing links with social cognition and ASD in the first paragraph of the revised manuscript (lines 46-56). In particular, integrating multisensory BM cues is foundational for perceiving and attending to other people and developing further social interaction. However, such ability is usually compromised in people with social deficits, such as individuals with autism spectrum disorder (ASD) (Feldman et al., 2018), and even in non-clinical populations with high autistic traits (Ujiie et al., 2015). These behavioral findings underline the close relationship between multisensory BM processing and one’s social cognitive capability, motivating us to further explore this issue at the neural level in the current study. We have also modified the relevant content in the last paragraph of the Introduction (lines 100-108), briefly mentioning the methods that we used to investigate this issue.

The use of terminology related to neural oscillations which are entraining to the BM seems to suggest that the rhythmic tracking inevitably stems from the shaping of existing intrinsic dynamics of the brain. I am not sure this is necessarily the case. I would therefore adopt a more concrete jargon for the description of the entrainment seen in this study. If a discussion over internal dynamics shaped by external stimuli should be invoked, it should be done explicitly with appropriate references (but in my opinion, it isn't quite required).

Please refer to our response to a similar point raised in the public review of Reviewer #2.

References

Bardi, L., Regolin, L., & Simion, F. (2014). The First Time Ever I Saw Your Feet: Inversion Effect in Newborns’ Sensitivity to Biological Motion. Developmental Psychology, 50. https://doi.org/10.1037/a0034678

Baron-Cohen, S., Wheelwright, S., Skinner, R., Martin, J., & Clubley, E. (2001). The autism-spectrum quotient (AQ): Evidence from Asperger syndrome/highfunctioning autism, males and females, scientists and mathematicians. Journal of Autism and Developmental Disorders, 31(1), 5–17. https://doi.org/10.1023/a:1005653411471

Batterink, L. (2020). Syllables in Sync Form a Link: Neural Phase-locking Reflects Word Knowledge during Language Learning. Journal of Cognitive Neuroscience, 32(9), 1735–1748. https://doi.org/10.1162/jocn_a_01581

Cirelli, L. K., Spinelli, C., Nozaradan, S., & Trainor, L. J. (2016). Measuring Neural Entrainment to Beat and Meter in Infants: Effects of Music Background. Frontiers in Neuroscience, 10. https://doi.org/10.3389/fnins.2016.00229

Ding, N., Melloni, L., Zhang, H., Tian, X., & Poeppel, D. (2016). Cortical tracking of hierarchical linguistic structures in connected speech. Nature Neuroscience, 19(1), 158–164. https://doi.org/10.1038/nn.4186

Duecker, K., Doelling, K. B., Breska, A., Coffey, E. B. J., Sivarao, D. V., & Zoefel, B. (2024). Challenges and approaches in the study of neural entrainment. Journal of Neuroscience, 44(40). https://doi.org/10.1523/JNEUROSCI.1234-24.2024

Falck-Ytter, T., Nyström, P., Gredebäck, G., Gliga, T., Bölte, S., & the EASE team. (2018). Reduced orienting to audiovisual synchrony in infancy predicts autism diagnosis at 3 years of age. Journal of Child Psychology and Psychiatry, 59(8), 872–880. https://doi.org/10.1111/jcpp.12863

Feldman, J. I., Dunham, K., Cassidy, M., Wallace, M. T., Liu, Y., & Woynaroski, T. G. (2018). Audiovisual multisensory integration in individuals with autism spectrum disorder: A systematic review and meta-analysis. Neuroscience & Biobehavioral Reviews, 95, 220–234. https://doi.org/10.1016/j.neubiorev.2018.09.020

Grossman, E. D., & Blake, R. (2001). Brain activity evoked by inverted and imagined biological motion. Vision Research, 41(10), 1475–1482. https://doi.org/10.1016/S0042-6989(00)00317-5

Henry, M. J., & Obleser, J. (2012). Frequency modulation entrains slow neural oscillations and optimizes human listening behavior. Proceedings of the National Academy of Sciences, 109(49), 20095–20100. https://doi.org/10.1073/pnas.1213390109

Herrmann, C. S., Murray, M. M., Ionta, S., Hutt, A., & Lefebvre, J. (2016). Shaping Intrinsic Neural Oscillations with Periodic Stimulation. Journal of Neuroscience, 36(19), 5328–5337. https://doi.org/10.1523/JNEUROSCI.0236-16.2016

Hosseinian, T., Yavari, F., Biagi, M. C., Kuo, M.-F., Ruffini, G., Nitsche, M. A., & Jamil, A. (2021). External induction and stabilization of brain oscillations in the human. Brain Stimulation, 14(3), 579–587. https://doi.org/10.1016/j.brs.2021.03.011

Klin, A., Lin, D. J., Gorrindo, P., Ramsay, G., & Jones, W. (2009). Two-year-olds with autism orient to non-social contingencies rather than biological motion. Nature, 459(7244), 257–261. https://doi.org/10.1038/nature07868

Laurienti, P. J., Perrault, T. J., Stanford, T. R., Wallace, M. T., & Stein, B. E. (2005). On the use of superadditivity as a metric for characterizing multisensory integration in functional neuroimaging studies. Experimental Brain Research, 166(3), 289–297. https://doi.org/10.1007/s00221-005-2370-2

Lenc, T., Keller, P. E., Varlet, M., & Nozaradan, S. (2018). Neural tracking of the musical beat is enhanced by low-frequency sounds. Proceedings of the National Academy of Sciences, 115(32), 8221–8226. https://doi.org/10.1073/pnas.1801421115

Metzger, B. A., Magnotti, J. F., Wang, Z., Nesbitt, E., Karas, P. J., Yoshor, D., & Beauchamp, M. S. (2020). Responses to Visual Speech in Human Posterior Superior Temporal Gyrus Examined with iEEG Deconvolution. The Journal of Neuroscience: The Official Journal of the Society for Neuroscience, 40(36), 6938–6948. https://doi.org/10.1523/JNEUROSCI.0279-20.2020

Moreau, C. N., Joanisse, M. F., Mulgrew, J., & Batterink, L. J. (2022). No statistical learning advantage in children over adults: Evidence from behaviour and neural entrainment. Developmental Cognitive Neuroscience, 57, 101154. https://doi.org/10.1016/j.dcn.2022.101154

Nozaradan, S., Peretz, I., & Mouraux, A. (2012). Selective Neuronal Entrainment to the Beat and Meter Embedded in a Musical Rhythm. Journal of Neuroscience, 32(49), 17572–17581. https://doi.org/10.1523/JNEUROSCI.3203-12.2012

Obleser, J., & Kayser, C. (2019). Neural Entrainment and Attentional Selection in the Listening Brain. Trends in Cognitive Sciences, 23(11), 913–926. https://doi.org/10.1016/j.tics.2019.08.004

Peter, V., Goswami, U., Burnham, D., & Kalashnikova, M. (2023). Impaired neural entrainment to low frequency amplitude modulations in English-speaking children with dyslexia or dyslexia and DLD. Brain and Language, 236, 105217. https://doi.org/10.1016/j.bandl.2022.105217

Shen, L., Lu, X., Wang, Y., & Jiang, Y. (2023). Audiovisual correspondence facilitates the visual search for biological motion. Psychonomic Bulletin & Review, 30(6), 2272–2281. https://doi.org/10.3758/s13423-023-02308-z

Shen, L., Lu, X., Yuan, X., Hu, R., Wang, Y., & Jiang, Y. (2023). Cortical encoding of rhythmic kinematic structures in biological motion. NeuroImage, 268, 119893. https://doi.org/10.1016/j.neuroimage.2023.119893

Simion, F., Regolin, L., & Bulf, H. (2008). A predisposition for biological motion in the newborn baby. Proceedings of the National Academy of Sciences, 105(2), 809–813. https://doi.org/10.1073/pnas.0707021105

Stanford, T. R., Quessy, S., & Stein, B. E. (2005). Evaluating the Operations Underlying Multisensory Integration in the Cat Superior Colliculus. Journal of Neuroscience, 25(28), 6499–6508. https://doi.org/10.1523/JNEUROSCI.5095-04.2005

Stevenson, R. A., Ghose, D., Fister, J. K., Sarko, D. K., Altieri, N. A., Nidiffer, A. R., Kurela, L. R., Siemann, J. K., James, T. W., & Wallace, M. T. (2014). Identifying and Quantifying Multisensory Integration: A Tutorial Review. Brain Topography, 27(6), 707–730. https://doi.org/10.1007/s10548-014-0365-7

Troje, N. F., & Westhoff, C. (2006). The Inversion Effect in Biological Motion Perception: Evidence for a “Life Detector”? Current Biology, 16(8), 821–824. https://doi.org/10.1016/j.cub.2006.03.022

Ujiie, Y., Asai, T., & Wakabayashi, A. (2015). The relationship between level of autistic traits and local bias in the context of the McGurk effect. Frontiers in Psychology, 6. https://doi.org/10.3389/fpsyg.2015.00891

Vallortigara, G., & Regolin, L. (2006). Gravity bias in the interpretation of biological motion by inexperienced chicks. Current Biology, 16(8), R279–R280. https://doi.org/10.1016/j.cub.2006.03.052

Vanrie, J., & Verfaillie, K. (2004). Perception of biological motion: A stimulus set of human point-light actions. Behavior Research Methods, Instruments, & Computers, 36(4), 625–629. https://doi.org/10.3758/BF03206542

Wang, L., & Jiang, Y. (2012). Life motion signals lengthen perceived temporal duration. Proceedings of the National Academy of Sciences of the United States of America, 109(11), E673-677. https://doi.org/10.1073/pnas.1115515109

Wang, L., Yang, X., Shi, J., & Jiang, Y. (2014). The feet have it: Local biological motion cues trigger reflexive attentional orienting in the brain. NeuroImage, 84, 217–224. https://doi.org/10.1016/j.neuroimage.2013.08.041

Wang, Y., Zhang, X., Wang, C., Huang, W., Xu, Q., Liu, D., Zhou, W., Chen, S., & Jiang, Y. (2022). Modulation of biological motion perception in humans by gravity. Nature Communications, 13(1), Article 1. https://doi.org/10.1038/s41467-022-30347-y

Wright, T. M., Pelphrey, K. A., Allison, T., McKeown, M. J., & McCarthy, G. (2003). Polysensory Interactions along Lateral Temporal Regions Evoked by Audiovisual Speech. Cerebral Cortex, 13(10), 1034–1043. https://doi.org/10.1093/cercor/13.10.1034

Version 2

Abstract

Abstract

When observing others’ behaviors, we continuously integrate their movements with the corresponding sounds to enhance perception and develop adaptive responses. However, how the human brain integrates these complex audiovisual cues based on their natural temporal correspondence remains unknown. Using electroencephalogram, we demonstrated that rhythmic cortical activity tracked the hierarchical rhythmic structures in audiovisually congruent human walking movements and footstep sounds. Remarkably, the cortical tracking effects at different time scales exhibit distinct modes of multisensory integration: an additive mode in a basic-level, narrower temporal integration window (step-cycle) and a super-additive enhancement in a higher-order, broader temporal window (gait-cycle). Moreover, only the cortical tracking of higher-order rhythmic structures is specialized for the multisensory integration of human motion signals and correlates with individuals’ autistic traits, suggesting its functional relevance to biological motion perception and social cognition. These findings unveil the multifaceted roles of entrained cortical activity in the multisensory perception of human motion, shedding light on how neural coding of hierarchical temporal structures orchestrates the processing of complex, rhythmic stimuli in natural contexts.

Introduction

The perception of biological motion (BM), the movements of living creatures, is a fundamental ability of the human visual system. Extensive evidence shows that humans can readily perceive BM from a visual display depicting just a handful of light dots attached to the head and major joints of a moving person (Blake & Shiffrar, 2007). Nevertheless, in real life, BM perception often occurs in multisensory contexts. For instance, one may simultaneously hear footstep sounds while seeing others walking. The integration of these visual and auditory BM cues facilitates the detection, discrimination, and attentional processing of BM (Mendonça et al., 2011; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2013; van der Zwan et al., 2009). Notably, such benefits are diminished when the visual BM is deprived of characteristic kinematic cues but not low-level motion attributes (Brooks et al., 2007; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2010), and the temporal windows of perceptual audiovisual synchrony are different between BM and non-BM stimuli (Arrighi et al., 2006; Saygin et al., 2008), highlighting the specificity of audiovisual BM processing. This specificity may relate to the evolutionary significance of BM and its relevance in social situations. In particular, integrating multisensory BM cues is foundational for perceiving and attending to other people and developing further social interaction. Such ability is usually compromised in people with social deficits, such as individuals with autism spectrum disorder (ASD) (Feldman et al., 2018), and even in non-clinical populations with high autistic traits (Ujiie et al., 2015). These findings underline the unique contribution of multisensory BM processing to human perception and social cognition. However, despite the behavioral evidence, the neural encoding of audiovisual BM cues and its possible link with individuals’ social cognitive capability remains largely unexplored.

An intrinsic property of human movements (such as walking and running) is that they are rhythmic and accompanied by frequency-congruent sounds. The audiovisual integration (AVI) of such rhythmic stimuli may involve cortical entrainment, a process that brain activity aligning with and tracking external rhythms, revealed by increased power or phase coherence of neural oscillations at corresponding frequencies (Bauer et al., 2020; Ding et al., 2016; Obleser & Kayser, 2019). Studies based on simple or discrete stimuli show that temporal congruency in auditory and visual rhythms significantly enhances the cortical tracking of rhythmic stimulations in both modalities (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). Unlike these stimuli, BM conveys complex hierarchical rhythmic structures corresponding to integration windows at multiple temporal scales. For example, the human locomotion movement has a narrower integration window consisting of each step (i.e., step cycle) and a broader integration window incorporating the opponent motion of the two feet (i.e., gait cycle). A recent study suggests that neural tracking of these nested kinematic structures contributes to the spatiotemporal integration of visual BM cues in different manners (Shen, Lu, Yuan, et al., 2023). However, it remains open whether and how the cortical tracking of hierarchical rhythmic structures underpins the AVI of BM information.

To tackle this issue, we recorded electroencephalogram (EEG) signals from participants who viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions in Experiments 1a & 1b (Fig. 1). An enhanced cortical tracking effect in the AV condition compared to each unisensory condition will indicate significant multisensory gains. Moreover, we adopted an additive model to classify multisensory integration based on the AV vs A+V comparison. This model assumes independence between inputs from each sensory modality and distinguishes among sub-additive (AV < A+V), additive (AV = A+V), and super-additive (AV > A+V) response modes (see a review by Stevenson et al., 2014). The additive mode represents a linear combination between two modalities. In contrast, the super-additive and sub-additive modes indicate non-linear interaction processing, either with potentiated neural activation to facilitate the perception or detection of near-threshold signals (super-additive) or a deactivation mechanism to minimize the processing of redundant information cross-modally (sub-additive) (Laurienti et al., 2005; Metzger et al., 2020; Stanford et al., 2005; Wright et al., 2003).

Illustrations of audiovisual stimuli and experimental procedures.

The illustration was based on stimuli with a gait-cycle frequency of 1 Hz. (a) Visual stimuli. The left panel depicts the static schematic of upright and inverted point-light walkers. The right panel shows the keyframes from a gait cycle of the BM sequence. The colors of dots and lines between dots are for illustration only and are not shown in the experiments. (b) Auditory stimuli. The auditory sequences contain periodic impulses of footstep sounds whose peak amplitudes occur around the points when the foot strikes the ground. The duration of two successive impulses defines the gait cycle of footstep sounds, which is temporally congruent (Con) or incongruent (InC) with the visual stimuli. (c) Experimental procedure and design. The color of the visual stimuli changed one or two times within 6 s in the catch trials but did not change in the experimental trials. Participants were required to report the number of changes when the point-light stimulus was replaced by a red fixation. In Experiment 1, participants viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions. The visual stimulus was the BM sequence in the V and AV conditions but a static frame from the sequence in the A condition. Experiment 2 included only the AV condition with different stimulus orientations (upright vs. inverted) and audiovisual congruency (congruent vs. incongruent).

Experiment 2 examined to what extent the AVI effect was specific to the multisensory processing of BM by using non-BM (inverted visual stimuli) as a control. Inversion disrupts the unique, gravity-compatible kinematic features of BM but not the rhythmic signals generated by low-level motion cues (Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Wang et al., 2022), thus is expected to interfere with the BM-specific neural processing. Participants perceived visual BM stimuli accompanied by temporally congruent or incongruent BM sounds. Comparing the congruency effect in neural responses between the upright and inverted conditions allowed us to verify whether the AVI of BM involves a mechanism distinct from that underlies the AVI of non-BM. In addition, to further explore the functional relevance of the BM-specific neural tracking effect, we examined its potential relationship with observers’ autistic traits. Previous behavioral studies found reduced orienting to audiovisually synchronized BM stimuli in ASD (Falck-Ytter et al., 2018; Klin et al., 2009). Since social cognitive deficits in ASD lie on a continuum extending from the clinical to nonclinical populations with different levels of autistic traits, as measured by the Autism-Spectrum Quotient (AQ) (Baron-Cohen et al., 2001), here we investigated the correlation between the cortical tracking of audiovisual BM and the individuals’ AQ scores.

Results

In all experiments, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed one or two times throughout the trial, and there was no color change in other trials. Participants were required to detect the color change of visual stimuli (0-2 times during one trial) to maintain attention. Behavioral analysis on all trials showed that their performances for the task were generally high and equally well in all conditions of Experiment 1a (mean accuracy > 98%; F (2, 46) = 0.814, p = .450, ƞp2 = 0.034), Experiment 1b (mean accuracy > 98%; F (2, 46) = 0.615, p = .545, ƞp2 = 0.026), and Experiment 2 (mean accuracy > 98%; F (3, 69) = 0.493, p = .688, ƞp2 = 0.021), indicating comparable attention state across conditions. The catch trials were excluded from the following EEG analysis.

Cortical tracking of rhythmic structures in audiovisual BM reveals AVI

Experiment 1a

In Experiment 1a, we examined the cortical tracking of rhythmic BM information under V, A, and AV conditions (Fig. 1c). We were interested in two critical rhythmic structures in the walking motion sequence, i.e., the gait cycle and the step cycle (Fig. 1a & 1b). During walking, each step of the left or right foot occurs alternatively to form a step cycle, and the antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, the frequency of a full gait cycle is 1 Hz, and the step-cycle frequency is 2 Hz. The strength of the cortical tracking effect was quantified by the amplitude peaks emerging from the EEG spectra at these frequencies.

As shown in the grand average amplitude spectra (Fig. 2a), both the responses in three conditions showed clear peaks at step-cycle frequency (2 Hz; V: t (23) = 6.963, p < 0.001; A: t (23) = 6.073, p < .001; AV: t (23) = 7.054, p < 0.001; FDR corrected). In contrast, at gait-cycle frequency (1 Hz), only the response to AV stimulation showed significant peaks (V: t (23) = -2.072, p = 0.975; A: t (23) = -0.054, p = 0.521; AV: t (23) = 4.059, p < 0.001; FDR corrected). Besides, we also observed significant peaks at 4 Hz in all three conditions (ps < 0.001, FDR corrected), which showed a similar audiovisual integration mode as 2 Hz (see more details in Supplementary Information).

Cortical tracking of visual (V), auditory (A), and audiovisual (AV) BM signals at gait-cycle and step-cycle frequencies.

(a) & (d) The amplitude spectra of EEG response in three conditions in Experiment 1a and Experiment 1b, respectively. The solid lines show the grand average amplitude over all electrodes and subjects. The shaded regions depict standard errors of the group mean. Asterisks indicate significant spectra peaks (one-sample t-test against zero; p < .05, FDR corrected). (b) & (e) The normalized amplitude at gait-cycle frequency in the AV condition exceeded the arithmetical sum of those in V and A conditions (AV > A+V), (c) & (f) but the normalized amplitude at step-cycle frequency in the AV condition was comparable to the sum of V and A (AV = A+V). Colored dots represent individual data in each condition. Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; m.s.: .05< p < .10; n.s.: p > .05.

Furthermore, we directly compared the cortical tracking effects between different conditions via a two-tailed paired t-test. At both 1 Hz (Fig. 2b) and 2 Hz (Fig. 2c), the amplitude in the AV condition was greater than that in the V condition (1 Hz: t (23) = 4.664, p < 0.001, Cohen’s d = 0.952; 2 Hz: t (23) = 5.132, p < 0.001, Cohen’s d = 1.048) and the A condition (1 Hz: t (23) = 2.391, p = 0.025, Cohen’s d = 0.488; 2 Hz: t (23) = 3.808, p < 0.001, Cohen’s d = 0.777), respectively, suggesting multisensory gains. More importantly, at 1Hz, the amplitude in the AV condition was significantly larger than the algebraic sum of those in the A and V conditions (t (23) = 3.028, p = 0.006, Cohen’s d = 0.618), indicating a super-additive audiovisual integration effect. While at 2Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.623, p = 0.539, Cohen’s d = -0.127), indicating additive audiovisual integration.

Experiment 1b

To further test whether such cortical tracking effect can apply to stimuli with a different speed, Experiment 1b altered the frequencies of the gait cycle and the corresponding step cycle to 0.83 Hz and 1.67 Hz while adopting the same paradigm as Experiment 1a. Consistent with Experiment 1a, the frequency-domain analysis revealed significant cortical tracking of the audiovisual stimuli at the new speeds. As shown in Fig. 2d, both the responses to V, A, and AV stimuli showed clear peaks at step-cycle frequency (1.67 Hz; V: t (23) = 3.473, p = .001; A: t (23) = 9.194, p < .001; AV: t (23) = 8.756, p < .001; FDR corrected) and its harmonics (3.33 Hz, ps < .001, FDR corrected). In contrast, at gait-cycle frequency (0.83 Hz), only the response to AV stimuli showed significant peaks (V: t (23) = -1.125, p = .846; A: t (23) = -2.449, p = .989; AV: t (23) = 3.052, p = .003; FDR corrected).

At both 0.83 Hz (Fig. 2e) and 1.67 Hz (Fig. 2f), the amplitude in the AV condition was stronger or marginally stronger than that in the V condition (0.83 Hz: t (23) = 2.665, p = .014, Cohen’s d = 0.544; 1.67 Hz: t (23) = 6.380, p < .001, Cohen’s d = 1.302) and the A condition (0.83 Hz: t (23) = 3.625, p < .001, Cohen’s d = 0.740; 1.67 Hz: t (23) = 1.752, p = .093, Cohen’s d = 0.358), respectively, suggesting multisensory gains. More importantly, at 0.83 Hz, the amplitude in the AV condition was significantly larger than the sum of those in the A and V conditions (t (23) = 3.240, p = .004, Cohen’s d = 0.661), indicating a super-additive audiovisual integration effect. While at 1.67 Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.735, p = .470, Cohen’s d = -0.150), indicating linear audiovisual integration. Significant peaks were also observed at 3.33 Hz in all three conditions (ps < 0.001, FDR corrected), which showed similar audiovisual integration mode as 1.67 Hz (see more details in Supplementary Information).

In summary, results from Experiments 1a & 1b consistently showed that the cortical tracking of the audiovisual signals at different temporal scales exhibit distinct audiovisual integration modes, i.e., the super-additive effect at gait-cycle frequency and the additive effect at step-cycle frequency, indicating that the cortical tracking effects at the two temporal scales might be driven by functionally dissociable mechanisms.

Cortical tracking of higher-order rhythmic structure contributes to the AVI of BM

To further explore whether and how the cortical tracking of rhythmic structures contributes to the specialized audiovisual processing of BM, both upright and inverted BM stimuli were adopted in Experiment 2. The task and the frequencies of visual stimuli in Experiment 2 were same as Experiment 1a. Specifically, participants were required to perform the change detection task when perceiving upright and inverted visual BM sequences (1 Hz for gait-cycle frequency and 2 Hz for step-cycle frequency) accompanied by frequency congruent (1 Hz) or incongruent (0.6 Hz and 1.4 Hz) footstep sounds (Fig. 1c). The audiovisual congruency effect, characterized by stronger neural responses in the audiovisual congruent condition compared with the incongruent condition, can be taken as an index of AVI (Fleming et al., 2020; Jones & Jarick, 2006; Maddox et al., 2015; Wuerger, Crocker-Buque, et al., 2012). A stronger congruency effect in the upright condition relative to the inverted condition characterizes an AVI process specific to BM information.

We calculated the audiovisual congruency effect for the upright and the inverted conditions, respectively. Then, we contrasted the congruency effect between the upright and inverted conditions to search for clusters showing a significant difference, which equaled identifying an interaction effect, using a cluster-based permutation test over all electrodes (n = 1000, alpha = 0.05; see Methods). At 1 Hz, the congruency effect in the upright condition was significantly stronger than that in the inverted condition in a cluster at the right hemisphere (Fig. 3a, lower panel, p = 0.029; C2, CPz, CP2, CP4, CP6, Pz, P2, P4, P6), revealing a BM-specific AVI process. Then we averaged the amplitude of electrodes within the significant cluster and further performed two-tailed paired t-tests to examine whether the congruency was significant in the upright and the inverted conditions, respectively. Results showed that (Fig. 3b) audiovisually congruent BM information enhanced the oscillatory amplitude relative to the incongruent ones only for upright BM stimuli (t (23) = 4.632, p < 0.001, Cohen’s d = 0.945) but not when visual BM was inverted (t (23) = 0.480, p = 0.635, Cohen’s d = 0.098).

Cortical tracking at gait-cycle rather than step-cycle frequency contributes to the BM-specific audiovisual integration (AVI) effect.

The lower panels in (a) and (d) depict the topographic maps of the BM-specific AVI effect, measured by the difference of congruency effect between the Upright and Inverted conditions at 1 Hz and 2 Hz, respectively. A significantly enhanced congruency effect in the upright condition relative to the inverted condition was observed at 1 Hz (marked by black dots) but not at 2 Hz. The amplitude at these significant electrodes was averaged to quantify the congruency effect for the upright and inverted conditions at 1 Hz (b) and 2 Hz (e). Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; n.s.: p > .05. Individuals’ autistic traits correlated with the BM-specific AVI effect at 1 Hz (c) but not at 2 Hz (f). Shaded regions indicate the 95% confidence intervals.

In contrast, at 2 Hz, no cluster showed a significantly different congruency effect between the upright and inverted conditions (Fig. 3d), suggesting no BM-specific AVI process. We then conducted further analysis based on the electrodes yielded by 1 Hz as marked in Fig. 3a. Results showed that both upright and inverted stimuli induced a significant congruency effect at 2 Hz (Fig. 3e; Upright: t (23) = 3.096, p = 0.005, Cohen’s d = 0.632; Inverted: t (23) = 2.672, p = 0.014, Cohen’s d = 0.545).

To verify the dissociation of the BM-specific AVI effects between the gait-cycle frequency and the step-cycle frequency, we performed a three-way repeated-measures ANOVA with cycle frequency (1 Hz vs. 2 Hz), orientation (upright vs. inverted), and audiovisual congruency (congruent vs. incongruent) as within-subject factors. We converted the amplitude data at each frequency condition into Z-scores to reduce the potential influence of amplitude magnitude difference between the two frequencies. The analysis of these data revealed a significant three-way interaction (F (1,23) = 7.501, p = 0.012, ƞp2 = 0.246), further supporting that the audiovisual integration processing of BM is different between 1 Hz and 2 Hz.

BM-specific cortical tracking correlates with autistic traits

Furthermore, we examined the link between individuals’ autistic traits and the neural responses underpinning the AVI of BM, measured by the difference of congruency effect between the upright and the inverted BM conditions, using Pearson correlation analysis. After removing one outlier (whose neural response exceeded 3 SD from the group mean), we observed an evident negative correlation between individuals’ AQ scores and their neural responses at 1 Hz (Fig. 3c, r = - 0.493, p = 0.017) but not at 2 Hz (Fig. 3f, r = -0.158, p = .460). The lack of significant results at 2 Hz was not attributable to electrode selection bias based on the significant cluster at 1 Hz, as similar results were observed when we performed the analysis on clusters showing non-selective significant congruency effects at 2 Hz (see the control analysis in Supplementary Information for details). Besides, we split the participants based on their median AQ score and found that, compared with the high AQ group, the low AQ group showed a greater BM-specific cortical tracking effect at 1 Hz but not at 2 Hz. These findings provide further support to the functional relevance between social cognition and cortical tracking of biological motion as well as its dissociation at the two temporal scales (see more details in Supplementary Information).

Discussion

The current study investigated the neural implementation for the AVI of human BM information and its functional implications. We found that, even under a motion-irrelevant color detection task, observers’ neural activity tracked the temporally corresponding audiovisual BM signals at the frequencies of two rhythmic structures, i.e., the higher-order structure of gait cycle at a larger integration window and the basic-level structure of step cycle at a smaller integration window. The strength of these cortical tracking effects was enhanced under the audiovisual condition than in the visual-only or auditory-only condition, indicating multisensory gains. More crucially, although the cortical tracking of both gait-cycle and step-cycle gain benefits from multisensory correspondence, the mechanisms underlying these two processes appear to be different. At step-cycle frequency, the cortical tracking effect in the AV condition equaled the additive sum of the unisensory conditions. Such linear integration may result from concurrent, independent processing of unisensory inputs without additional interaction of them (Stein et al., 2009). In contrast, at gait-cycle frequency, the congruent audiovisual signals led to a super-additive multisensory enhancement over the linear combination of auditory and visual conditions (AV > A+V), despite that there was no evident cortical tracking effect in the visual condition, different from previous findings obtained with a motion-relevant change detection task (Shen, Lu, Yuan, et al., 2023). This super-additive multisensory enhancement may bring about decreased thresholds of detection and identification (Stanford et al., 2005), allowing us to achieve a more clear and stable perception of the external environment and detect weak stimulus changes in time and respond adaptively.

Furthermore, results from Experiment 2 demonstrated that the cortical tracking of rhythmic structure corresponding to the gait cycle rather than the step cycle is relevant to the specialized processing of audiovisual BM information. In particular, the AVI effect at step-cycle frequency was significant for both upright and inverted BM signals and comparable between the two conditions, while the AVI effect at gait-cycle frequency was only significant in the upright condition and was greater than that in the inverted condition. The inversion effect has long been regarded as a marker of the specificity of BM processing in numerous behavioral and neuroimaging studies (Grossman & Blake, 2001; Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Vallortigara & Regolin, 2006; Wang et al., 2014; Wang & Jiang, 2012; Wang et al., 2022). Our current findings of the inversion effect in the cortical tracking of audiovisual BM at the gait-cycle suggest that the neural encoding of the higher-order rhythmic structure reflects the AVI of BM and contributes to the specialized processing of BM information. In contrast, the cortical tracking of the step-cycle may reflect the integration of basic motion signals and corresponding sounds. Together, these results reveal that the neural tracking of different levels of kinematic structures plays distinct roles in the AVI of BM, which may result from the interplay of stimulus-driven and domain-specific mechanisms. In addition, a recent study demonstrated that listening to frequency-congruent footstep sounds, compared with incongruent sounds, enhanced the visual search for human walkers but not for non-BM stimuli containing the same rhythmic signals, indicating that audiovisual correspondence specifically enhances the perceptual and attentional processing of BM (Shen, Lu, Wang, et al., 2023). Future research could examine whether the cortical tracking of rhythmic structures plays a functional role in this process, which may shed more light on the behavioral relevance of the cortical tracking effect to BM perception.

Besides the temporal dynamics of neural activity revealed by the cortical tracking process, we found that the BM-specific AVI effect was associated with neural activity in the right temporoparietal electrodes. This finding likely relates to the activation of the right posterior superior temporal sulcus (pSTS), a region responding to both auditory and visual BM information and being causally involved in BM perception (Bidet-Caulet et al., 2005; Grossman et al., 2005; Wang et al., 2022). While previous fMRI studies have observed STS activation when processing spatial or semantic correspondence between audiovisual BM (Meyer et al., 2011; Wuerger, Parkes, et al., 2012), whether this region also engages in the audiovisual processing of BM signals based on temporal correspondence remains unknown. The current study provides preliminary evidence for such a possibility, inviting future research to localize the exact source of the multisensory integration processes based on imaging data with high spatial and temporal resolutions, such as MEG.

Cortical tracking of external rhythms is also described as cortical entrainment in a broad sense (Ding et al., 2016; Obleser & Kayser, 2019). Controversy remains regarding the involvements of endogenous neural oscillations and stimulus-evoked responses in these processes (Duecker et al., 2024), as it is challenging to fully dissociate these components due to their intricate interplay (Herrmann et al., 2016; Hosseinian et al., 2021). Previous research has demonstrated that cortical tracking or entrainment plays a role in the multisensory processing of simple or discrete rhythmic signals (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). These findings may partially explain the non-selective AVI effect at the step-cycle in the current study. However, we found that the cortical tracking of the higher-order rhythmic structure formed by spatiotemporal integration of meaningful BM information (i.e., the gait cycle of upright walkers rather than inverted walkers) is selectively engaged by the AVI of BM, suggesting that the multisensory processing of natural continuous stimuli may involve unique mechanisms besides the purely stimulus-driven AVI process. These findings provide significant implications for the neural processing of natural audiovisual stimuli with complex temporal structures. Similar to BM, other natural rhythmic stimuli, like speech and music, also convey hierarchical structures that can entrain neural oscillations at different temporal scales, both in unisensory (Ding et al., 2016; Doelling & Poeppel, 2015) and multisensory contexts (Biau et al., 2022; Crosse et al., 2015; Nozaradan et al., 2016). Possibly, the audiovisual processing of these stimuli also engages multi-scale neural coding mechanisms that play distinct functions in perception. Investigating this issue and comparing the results with BM studies will help complete the picture of how the human brain integrates complex, rhythmic information sampled from different sensory modalities to orchestrate perception in a natural scenario.

Last but not least, our study demonstrated that the selective cortical tracking of higher-level rhythmic structure in audiovisually congruent BM signals negatively correlated with individual autistic traits. This finding highlights the functional significance of the neural tracking of audiovisual BM signals in social cognition. It also offers the first evidence that differences in audiovisual BM processing are already present in nonclinical individuals and associated with their autistic traits, beyond previous evidence for atypical audiovisual BM processing in ASD populations (Falck-Ytter et al., 2013, 2018; Klin et al., 2009), lending support to the continuum view of ASD (Baron-Cohen et al., 2001). Meanwhile, given that impaired audiovisual BM processing at the early stage may influence social development and result in cascading consequences for lifetime impairments in social interaction (Falck-Ytter et al., 2018; Klin et al., 2005), it is worth exploring neural tracking of audiovisual BM signals in children with different autistic levels, which may help reveal whether deficits in such ability could serve as an early neural hallmark for ASD.

Materials and methods

Participants

Seventy-two participants (mean age ± SD = 22.4 ± 2.6 years, 35 females) took part in the study, 24 for each of Experiment 1a, Experiment 1b, and Experiment 2. All of them had normal or corrected-to-normal vision and reported no history of neurological, psychiatric, or hearing disorders. They were naïve to the purpose of the study and gave informed consent according to procedures and protocols approved by the institutional review board of the Institute of Psychology, Chinese Academy of Sciences.

Stimuli

Visual stimuli

The visual stimuli (Fig. 1a, left panel) consisted of 13 point-light dots attached to the head and major joints of a human walker (Vanrie & Verfaillie, 2004). The point-light walker was presented at the center of the screen without translational motion. It conveys rhythmic structures specified by recurrent forward motions of bilateral limbs (Fig. 1a, right panel). Each step, regardless of left or right foot, occurs recurrently to form a step cycle. The antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, a full gait cycle took 1 second and was repeated 6 times to form a 6-second walking sequence. That is, the gait-cycle frequency is 1 Hz and the step-cycle frequency is 2 Hz. In Experiment 1b, the gait-cycle frequency was 0.83 Hz and the step-cycle frequency was 1.67 Hz. The gait cycle was repeated 6 times to form a 7.2-second walking sequence. The stimuli in Experiment 2 were the same as that in Experiment 1a. Meanwhile, the point-light BM was mirror-flipped vertically to generate inverted BM (Fig. 1a, left panel), which preserves the temporal structure of the stimuli but distorts its distinctive kinematic features, such as movement that is compatible with the effect of gravity (Shen, Lu, Yuan, et al., 2023; Troje & Westhoff, 2006; Wang et al., 2022).

Auditory stimuli

Auditory stimuli were continuous footstep sounds (6 s) with a sampling rate of 44,100 Hz. As shown in Fig. 1b, in Experiments 1a & 2, the gait-cycle frequency of congruent sounds was 1 Hz, which consisted of two steps or two impulses generated by each foot striking the ground within one gait cycle. The incongruent sounds included a faster (1.4 Hz) and a slower (0.60 Hz) sound. Both congruent and incongruent sounds were generated by manipulating the temporal interval between two successive impulses based on the same auditory stimuli. In Experiment 1b, the gait-cycle frequency of sound was 0.83 Hz.

Stimuli presentation

The visual stimuli were rendered white against a grey background and displayed on a CRT (cathode ray tube) monitor. Participants sat 60 cm from the computer screen (1280×1024 at 60 Hz; High: 37.5 cm; Width: 30 cm), with their heads held stationary on a chinrest. The auditory stimuli were presented binaurally over insert earphones. All stimuli were generated and presented using MATLAB together with the Psychophysics Toolbox (Brainard, 1997; Pelli, 1997).

Procedure and task

Experiment 1a

The experiment was conducted in an acoustically dampened and electromagnetically shielded chamber. Participants completed the task under three conditions (Visual: V; Auditory: A; Audiovisual: AV) with the same procedure (Fig. 1c) except for the stimuli. In the V condition, each trial began with a white fixation cross (0.42° × 0.42°) displayed at the center of a gray background for a random duration (0.8 s to 1 s). Subsequently, a 6-s point-light walker (3.05°×5.47°) walked toward the left or right at a constant walking cycle frequency (1 Hz). To maintain observers’ attention, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed (the RGB values changed from [255 255 255] to [207 207 207]) one or two times throughout the trial. Each change lasted 0.5 s. Observers were required to report the number of changes (0, 1, or 2) via keypresses as accurately as possible after the point-light display was replaced by a red fixation. The next trial started 2–3 s after the response. In the A condition, the 6 s-stimuli were replaced by a visually static BM figure accompanied by continuous footstep sounds. The frequency of footstep sounds was congruent with the frequency of visual BM in the V condition. In the AV condition, the stimuli were temporally congruent visual BM sequences (as in the V condition) and footstep sounds (as in the A condition). Three conditions were conducted in separate blocks. V condition was performed in the middle of A and AV conditions. The order of A and AV conditions was counterbalanced across participants. Each participant completed 40 experimental trials without changes and 10-15 catch trials in each condition, resulting in a total of 150-165 trials. In each condition, participants completed a practice session with 3 trials to get familiar with the task before the formal EEG experiment.

Experiment 1b

The procedure of Experiment 1b was the same as that for Experiment 1a but with two exceptions. First, to test if the cortical tracking effect can apply to stimuli with a different speed, we altered the frequencies of gait and step cycles to 0.83 Hz and 1.67 Hz. Second, we presented the 3 conditions (V, A, and AV) in a completely random order to eliminate the influence of presentation order. To minimize the potential influence of condition switch, we increased the trial number in the practice session from 3 to 14 for each condition.

Experiment 2

The procedure in Experiment 2 was similar to the AV condition in Experiment 1a, except that the visually displayed BM was accompanied by frequency congruent (1 Hz) or incongruent (0.6 or 1.4 Hz) footstep sounds. Each participant completed a total of 76 experiment trials, consisting of 36 congruent-trials, 20 incongruent-trials with faster sounds (1.4 Hz), and 20 incongruent-trials with slower sounds (0.6 Hz). These trials were assigned to 3 blocks based on the frequency of the footstep sounds, with the order of the three frequencies balanced across participants. Besides, an inverted BM was used as a control to investigate whether there is a specialized mechanism tuned to the AVI of life motion signals. The order of upright and inverted conditions was balanced across participants. Meanwhile, we measured the participants’ autistic traits by using the Autism-Spectrum Quotient, or AQ questionnaire (Baron-Cohen et al., 2001). Higher AQ scores indicate a higher level of autistic traits.

EEG recording and analysis

EEG was recorded at 1000 Hz using a SynAmps2 NeuroScan amplifier System with 64 electrodes placed on the scalp according to the international 10-20 system. Horizontal and vertical eye movements were measured via four additional electrodes placed on the outer canthus of each eye and the inferior and superior areas of the left orbit. Impedances were kept below 5 kΩ for all electrodes.

Preprocessing

The catch trials were excluded from EEG analysis. All preprocessing and further analyses were performed using the FieldTrip toolbox (Oostenveld et al., 2011; http://fieldtriptoolbox.org) in the MATLAB environment. EEG recordings were pass-filtered between 0.1 and 30 Hz, and down-sampled to 100 Hz. Then the continuous EEG data were cut into epochs ranging from -1s to 6 gait cycles (7.2 s in Experiment 1b and 6 s in other experiments) time-locked to the onset of the visual point-light stimuli. The epochs were visually inspected, and trials contaminated with excessive noise were excluded from the analysis. After the trial rejection, eye and cardiac artifacts were removed via independent component analysis based on the Runica algorithm (Bell & Sejnowski, 1995; Jung et al., 2000; Makeig, 2002). Then the cleaned data were re-referenced to the average mastoids (M1 and M2). To minimize the influence of stimulus-onset evoked activity on EEG spectral decomposition, the EEG recording before the onset of the stimulus and the first cycle (1 s in Experiments 1a & 2; 1.2 s in Experiment 1b) of each trial was excluded (Nozaradan et al., 2012a). After that, the EEG epochs were averaged across trials for each participant and condition.

Frequency-Domain analysis and statistics

A Fast Fourier Transform (FFT) with zero padding (1200) was used to convert the averaged EEG signals from the temporal domain to the spectral domain, resulting in a frequency resolution of 0.083 Hz, i.e., 1/12 Hz, which is sufficient for observing neural responses around the frequency of the rhythmic BM structures in all experiments. When performing FFT, a Hanning window was adopted to minimize spectral leakage. Then, to remove the 1/f trend of the response amplitude spectrum and identify spectral peaks, the response amplitude at each frequency was normalized by subtracting the average amplitude measured at the neighboring frequency bins (two bins on each side) (Nozaradan et al., 2012a). We calculated the normalized amplitude separately for each electrode (except for electrooculogram electrodes, CB1, and CB2), participant, and condition.

In Experiment 1, the normalized amplitude in all electrodes was averaged and a right-tailed one-sample t-test against zero was performed on the grand average amplitude to test whether the neural response in each frequency bin showed a significant tracking effect or spectral peak. This test was applied to all frequency bins below 5.33 Hz and multiple comparisons were controlled by false discovery rate (FDR) correction at p < 0.05 (Benjamini & Hochberg, 1995). In Experiment 2, to further identify the BM-specific AVI process, the audiovisual congruency effect was compared between the upright and inverted conditions using a cluster-based permutation test over all electrodes (1000 iterations, requiring a cluster size of at least 2 significant neighbors, a two-sided t-test at p < 0.05 on the clustered data) (Oostenveld et al., 2011; http://fieldtriptoolbox.org). This allowed us to identify the spatial distribution of the BM-specific congruency effect.

eLife Assessment

Wang et al. presented visual (dot) motion and/or the sound of a walking person and found solid evidence that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm, with some demonstration that the gait rhythm is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition). The valuable findings will be of wide interest to those examining biological motion perception and oscillatory processes more broadly. Some of the theoretical interpretations concerning entrainment must remain speculative when the authors cannot dissociate evoked responses from entrained oscillatory effects

Reviewer #1 (Public review):

Summary:

Shen et al. conducted three experiments to study the cortical tracking of the natural rhythms involved in biological motion (BM), and whether these involve audiovisual integration (AVI). They presented participants with visual (dot) motion and/or the sound of a walking person. They found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm. The gait rhythm specifically is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition, Experiments 1a/b), which is independent of the specific step frequency (Experiment 1b). Furthermore, audiovisual integration during tracking of gait was specific to BM, as it was absent (that is, the audiovisual congruency effect) when the walking dot motion was vertically inverted (Experiment 2). Finally, the study shows that an individual's autistic traits are negatively correlated with the BM-AVI congruency effect.

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

Weaknesses:

There is a concern of double-dipping in one of the tests (Experiment 2, Figure 3: interaction of Upright/Inverted X Congruent/Incongruent). I raised this concern on the original submission, and it has not been resolved properly. The follow-up statistical test (after channel selection using the interaction contrast permutation test) still is geared towards that same contrast, even though the latter is now being tested differently. (Perhaps not explicitly testing the interaction, but in essence still testing the same.) A very simple solution would be to remove the post-hoc statistical tests and simply acknowledge that you're comparing simple means, while the statistical assessment was already taken care of using the permutation test. (In other words: the data appear compelling because of the cluster test, but NOT because of the subsequent t-tests.)

Reviewer #2 (Public review):

Summary:

The authors evaluate spectral changes in electroencephalography (EEG) data as a function of the congruency of audio and visual information associated with biological motion (BM) or non-biological motion. The results show supra-additive power gains in the neural response to gait dynamics, with trials in which audio and visual information was presented simultaneously producing higher average amplitude than the combined average power for auditory and visual conditions alone. Further analyses suggest that such supra-additivity is specific to BM and emerges from temporoparietal areas. The authors also find that the BM-specific supra-additivity is negatively correlated with autism traits.

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

Weaknesses:

In the revised version of the paper, the manuscript better relays the results and anticipates analyses, and this version adequately resolves some concerns I had about analysis details. Still, it is my view that the findings of the study are basic neural correlate results that do not provide insights into neural mechanisms or the causal relevance of neural effects towards behavior and cognition. The presence of an inversion effect suggests that the supra-additivity is related to cognition, but that leaves open whether any detected neural pattern is actually consequential for multi-sensory integration (i.e., correlation is not causation). In other words, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated in the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion, and it is possible that such neural correlates relate to behaviorally relevant neural mechanisms, but based on the current task and associated analyses this has not been shown.

Author response:

The following is the authors’ response to the previous reviews.

Public Reviews:

Reviewer #1 (Public review):

Summary:

Shen et al. conducted three experiments to study the cortical tracking of the natural rhythms involved in biological motion (BM), and whether these involve audiovisual integration (AVI). They presented participants with visual (dot) motion and/or the sound of a walking person. They found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm. The gait rhythm specifically is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition, Experiments 1a/b), which is independent of the specific step frequency (Experiment 1b). Furthermore, audiovisual integration during tracking of gait was specific to BM, as it was absent (that is, the audiovisual congruency effect) when the walking dot motion was vertically inverted (Experiment 2). Finally, the study shows that an individual's autistic traits are negatively correlated with the BM-AVI congruency effect.

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

Weaknesses:

There is a concern of double-dipping in one of the tests (Experiment 2, Figure 3: interaction of Upright/Inverted X Congruent/Incongruent). I raised this concern on the original submission, and it has not been resolved properly. The follow-up statistical test (after channel selection using the interaction contrast permutation test) still is geared towards that same contrast, even though the latter is now being tested differently. (Perhaps not explicitly testing the interaction, but in essence still testing the same.) A very simple solution would be to remove the post-hoc statistical tests and simply acknowledge that you're comparing simple means, while the statistical assessment was already taken care of using the permutation test. (In other words: the data appear compelling because of the cluster test, but NOT because of the subsequent t-tests.)

We are sorry that we did not explain this issue clearly before, which might have caused some misunderstanding. When performing the cluster-based permutation test, we only tested whether the audiovisual congruency effect (congruent vs. incongruent) between the upright and inverted conditions was significantly different [i.e., (UprCon – UprInc) vs. (InvCon – InvInc)], without conducting extra statistical analyses on whether the congruency effect was significant in each orientation condition. Such an analysis yielded a cluster with a significant interaction between audiovisual integration and BM orientation for the cortical tracking effect at 1Hz (but not at 2Hz). However, this does not provide valid information about whether the audiovisual congruency effect at this cluster is significant in each orientation condition, given that a significant interaction effect may result from various patterns of data across conditions: such as significant congruency effects in both orientation conditions (Author response image 1a), a significant congruency effect in the upright condition and a non-significant effect in the inverted condition (Author response image 1b), or even non-significant yet opposite effects in the two conditions (Author response image 1c). Here, our results conform to the second pattern, indicating that cortical tracking of the high-order gait cycles involves a domain-specific process exclusively engaged in the AVI of BM. In a similar vein, the non-significant interaction found at 2Hz does not necessarily indicate that the congruency effect is non-significant in each orientation condition (Author response image 1f&e). Indeed, the congruency effect was significant in both the upright and inverted conditions at 2Hz in our study despite the non-significant interaction, suggesting that neural tracking of the lower-order step cycles is associated with a domain-general AVI process mostly driven by temporal correspondence in physical stimuli.

Therefore, we need to perform subsequent t-tests to examine the significance of the simple effects in the two orientation conditions, which do not duplicate the clusterbased permutation test (for interaction only) and cause no double-dipping. Results from interaction and simple effects, put together, provide solid evidence that the cortical tracking of higher-order and lower-order rhythms involves BM-specific and domaingeneral audiovisual processing, respectively.

To avoid ambiguity, we have removed the sentence “We calculated the audiovisual congruency effect for the upright and the inverted conditions” (line 194, which referred to the calculation of the indices rather than any statistical tests) from the manuscript. We have also clarified the meanings of the findings based on the interaction and simple effects together at the two temporal scales, respectively (Lines 205-207; Lines 213-215).

Examples of different patterns of data yielding a significant or nonsignificant interaction effect.

Reviewer #2 (Public review):

Summary:

The authors evaluate spectral changes in electroencephalography (EEG) data as a function of the congruency of audio and visual information associated with biological motion (BM) or non-biological motion. The results show supra-additive power gains in the neural response to gait dynamics, with trials in which audio and visual information was presented simultaneously producing higher average amplitude than the combined average power for auditory and visual conditions alone. Further analyses suggest that such supra-additivity is specific to BM and emerges from temporoparietal areas. The authors also find that the BM-specific supra-additivity is negatively correlated with autism traits.

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

Weaknesses:

In the revised version of the paper, the manuscript better relays the results and anticipates analyses, and this version adequately resolves some concerns I had about analysis details. Still, it is my view that the findings of the study are basic neural correlate results that do not provide insights into neural mechanisms or the causal relevance of neural effects towards behavior and cognition. The presence of an inversion effect suggests that the supra-additivity is related to cognition, but that leaves open whether any detected neural pattern is actually consequential for multi-sensory integration (i.e., correlation is not causation). In other words, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated in the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion, and it is possible that such neural correlates relate to behaviorally relevant neural mechanisms, but based on the current task and associated analyses this has not been shown.

Thank you for providing these thoughtful comments regarding the theoretical implications of our neural findings. Previous behavioral evidence highlights the specificity of the audiovisual integration (AVI) of biological motion (BM) and reveals the impairment of such ability in individuals with autism spectrum disorder. However, the neural implementation underlying the AVI of BM, its specificity, and its association with autistic traits remain largely unknown. The current study aimed to address these issues.

It is noteworthy that the operation of multisensory integration does not always depend on specific tasks, as our brains tend to integrate signals from different sensory modalities even when there is no explicit task. Hence, many studies have investigated multisensory integration at the neural level without examining its correlation with behavioral performance. For example, the widely known super-additivity mode for multisensory integration proposed by Perrault and colleagues was based on single-cell recording findings without behavioral tasks (Perrault et al., 2003, 2005). As we mentioned in the manuscript, the super-additive and sub-additive modes indicate non-linear interaction processing, either with potentiated neural activation to facilitate the perception or detection of near-threshold signals (super-additive) or a deactivation mechanism to minimize the processing of redundant information cross-modally (subadditive) (Laurienti et al., 2005; Metzger et al., 2020; Stanford et al., 2005; Wright et al., 2003). Meanwhile, the additive integration mode represents a linear combination between two modalities. Distinguishing among these integration modes helps elucidate the neural mechanism underlying AVI in specific contexts, even though sometimes, the neural-level AVI effects do not directly correspond to a significant behavioral-level AVI effect (Ahmed et al., 2023; Metzger et al., 2020). In the current study, we unveiled the dissociation of multisensory integration modes between neural responses at two temporal scales (Exps. 1a & 1b), which may involve the cooperation of a domain-specific and a domain-general AVI processes (Exp. 2). While these findings were not expected to be captured by a single behavioral index, they revealed the multifaceted mechanism whereby hierarchical cortical activity supports audiovisual BM integration. They also advance our understanding of the emerging view that multi-timescale neural dynamics coordinate multisensory integration (Senkowski & Engel, 2024), especially from the perspective of natural stimuli processing.

Meanwhile, our finding that the cortical tracking of higher-order rhythmic structure in audiovisual BM specifically correlated with individual autistic traits extends previous behavioral evidence that ASD children exhibited reduced orienting to audiovisual synchrony in BM (Falck-Ytter et al., 2018), offering new evidence that individual differences in audiovisual BM processing are present at the neural level and associated with autistic traits. This finding opens the possibility of utilizing the cortical tracking of BM as a potential neural maker to assist the diagnosis of autism spectrum disorder (see more details in our Discussion Lines 334-346).

However, despite the main objective of the current study focusing on the neural processing of BM, we agree with the reviewer that it would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to BM perception, for further justifying that inputs are being integrated in the service of behavior. In the current study, we adopted a color-change detection task entirely unrelated to audiovisual correspondence but only for maintaining participants’ attention. The advantage of this design is that it allows us to investigate whether and how the human brain integrates audiovisual BM information under task-irrelevant settings, as people in daily life can integrate such information even without a relevant task. However, this advantage is accompanied by a limitation: the task does not facilitate the direct examination of the correlation between neural responses and behavioral performance, since the task performance was generally high (mean accuracy >98% in all experiments). Future research could investigate this issue by introducing behavioral tasks more relevant to BM perception (e.g., Shen et al., 2023). They could also apply advanced neuromodulation techniques to elucidate the causal relevance of the cortical tracking effect to behavior (e.g., Ko sem et al., 2018, 2020).

We have discussed the abovementioned points as a separate paragraph in the revised manuscript (Lines 322-333). In addition, since the scope of the current study does not involve a causal correlation with behavioral performance, we have removed or modified the descriptions related to "functional relevance" in the manuscript (Abstract; Introduction, lines 101-103; Results, lines 239; Discussion, line 336; Supplementary Information, line 794、803). Moreover, we have strengthened the descriptions of the theoretical implications of the current findings in the abstract.

We hope these changes adequately address your concern.

References

Ahmed, F., Nidiffer, A. R., O’Sullivan, A. E., Zuk, N. J., & Lalor, E. C. (2023). The integration of continuous audio and visual speech in a cocktail-party environment depends on attention. Neuroimage, 274, 120143. https://doi.org/10.1016/j.neuroimage.2023.120143

Falck-Ytter, T., Nystro m, P., Gredeba ck, G., Gliga, T., Bo lte, S., & the EASE team. (2018). Reduced orienting to audiovisual synchrony in infancy predicts autism diagnosis at 3 years of age. Journal of Child Psychology and Psychiatry, 59(8), 872–880. https://doi.org/10.1111/jcpp.12863

Ko sem, A., Bosker, H., Jensen, O., Hagoort, P., & Riecke, L. (2020). Biasing the Perception of Spoken Words with Transcranial Alternating Current Stimulation. Journal of Cognitive Neuroscience, 32, 1–10. https://doi.org/10.1162/jocn_a_01579

Ko sem, A., Bosker, H. R., Takashima, A., Meyer, A., Jensen, O., & Hagoort, P. (2018). Neural Entrainment Determines the Words We Hear. Current Biology, 28(18), 2867-2875.e3. https://doi.org/10.1016/j.cub.2018.07.023

Laurienti, P. J., Perrault, T. J., Stanford, T. R., Wallace, M. T., & Stein, B. E. (2005). On the use of superadditivity as a metric for characterizing multisensory integration in functional neuroimaging studies. Experimental Brain Research, 166(3), 289–297. https://doi.org/10.1007/s00221-005-2370-2

Metzger, B. A., Magnotti, J. F., Wang, Z., Nesbitt, E., Karas, P. J., Yoshor, D., & Beauchamp, M. S. (2020). Responses to Visual Speech in Human Posterior Superior Temporal Gyrus Examined with iEEG Deconvolution. The Journal of Neuroscience: The Official Journal of the Society for Neuroscience, 40(36), 6938–6948. https://doi.org/10.1523/JNEUROSCI.0279-20.2020

Perrault, T. J., Vaughan, J. W., Stein, B. E., & Wallace, M. T. (2003). Neuron-Specific Response Characteristics Predict the Magnitude of Multisensory Integration. Journal of Neurophysiology, 90(6), 4022–4026. https://doi.org/10.1152/jn.00494.2003

Perrault, T. J., Vaughan, J. W., Stein, B. E., & Wallace, M. T. (2005). Superior Colliculus Neurons Use Distinct Operational Modes in the Integration of Multisensory Stimuli. Journal of Neurophysiology, 93(5), 2575–2586. https://doi.org/10.1152/jn.00926.2004

Senkowski, D., & Engel, A. K. (2024). Multi-timescale neural dynamics for multisensory integration. Nature Reviews Neuroscience, 25(9), 625–642. https://doi.org/10.1038/s41583-024-00845-7

Shen, L., Lu, X., Wang, Y., & Jiang, Y. (2023). Audiovisual correspondence facilitates the visual search for biological motion. Psychonomic Bulletin & Review, 30(6), 2272–2281. https://doi.org/10.3758/s13423-023-02308-z

Stanford, T. R., Quessy, S., & Stein, B. E. (2005). Evaluating the Operations Underlying Multisensory Integration in the Cat Superior Colliculus. Journal of Neuroscience, 25(28), 6499–6508. https://doi.org/10.1523/JNEUROSCI.5095-04.2005

Wright, T. M., Pelphrey, K. A., Allison, T., McKeown, M. J., & McCarthy, G. (2003). Polysensory Interactions along Lateral Temporal Regions Evoked by Audiovisual Speech. Cerebral Cortex, 13(10), 1034–1043. https://doi.org/10.1093/cercor/13.10.1034

Version 3

Abstract

Abstract

When observing others’ behaviors, we continuously integrate their movements with the corresponding sounds to enhance perception and develop adaptive responses. However, how the human brain integrates these complex audiovisual cues based on their natural temporal correspondence remains unclear. Using electroencephalogram (EEG), we demonstrated that rhythmic cortical activity tracked the hierarchical rhythmic structures in audiovisually congruent human walking movements and footstep sounds. Remarkably, the cortical tracking effects exhibit distinct multisensory integration modes at two temporal scales: an additive mode in a lower-order, narrower temporal integration window (step-cycle) and a super-additive enhancement in a higher-order, broader temporal window (gait-cycle). Furthermore, while neural responses at the lower-order timescale reflect a domain-general audiovisual integration process, cortical tracking at the higher-order timescale is exclusively engaged in the integration of biological motion cues. In addition, only this higher-order, domain-specific cortical tracking effect correlates with individuals’ autistic traits, highlighting its potential as a neural marker for autism spectrum disorder. These findings unveil the multifaceted mechanism whereby rhythmic cortical activity supports the multisensory integration of human motion, shedding light on how neural coding of hierarchical temporal structures orchestrates the processing of complex, natural stimuli across multiple timescales.

Introduction

The perception of biological motion (BM), the movements of living creatures, is a fundamental ability of the human visual system. Extensive evidence shows that humans can readily perceive BM from a visual display depicting just a handful of light dots attached to the head and major joints of a moving person (Blake & Shiffrar, 2007). Nevertheless, in real life, BM perception often occurs in multisensory contexts. For instance, one may simultaneously hear footstep sounds while seeing others walking. The integration of these visual and auditory BM cues facilitates the detection, discrimination, and attentional processing of BM (Mendonça et al., 2011; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2013; van der Zwan et al., 2009). Notably, such benefits are diminished when the visual BM is deprived of characteristic kinematic cues but not low-level motion attributes (Brooks et al., 2007; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2010), and the temporal windows of perceptual audiovisual synchrony are different between BM and non-BM stimuli (Arrighi et al., 2006; Saygin et al., 2008), highlighting the specificity of audiovisual BM processing. This specificity may relate to the evolutionary significance of BM and its pivotal role in social situations. In particular, integrating multisensory BM cues is foundational for perceiving and attending to other people and developing further social interaction. Such ability is usually compromised in people with social deficits, such as individuals with autism spectrum disorder (ASD) (Falck-Ytter et al., 2013, 2018; Feldman et al., 2018). These findings underline the unique contribution of multisensory BM processing to human perception and social cognition. However, despite the behavioral evidence, the neural coding of audiovisual BM cues and its possible link with individuals’ social cognitive capability remains largely unexplored.

An intrinsic property of human movements (such as walking and running) is that they are rhythmic and accompanied by frequency-congruent sounds. The audiovisual integration (AVI) of such rhythmic stimuli may involve a process that brain activity aligning with and tracking external rhythms, revealed by increased power or phase coherence of neural oscillations at corresponding frequencies (Bauer et al., 2020; Ding et al., 2016; Obleser & Kayser, 2019). Studies based on simple or discrete stimuli show that temporal congruency in auditory and visual rhythms significantly enhances the cortical tracking of rhythmic stimulations in both modalities (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). Unlike these stimuli, BM conveys complex hierarchical rhythmic structures corresponding to integration windows at multiple temporal scales. For example, the human locomotion movement has a narrower integration window consisting of each step (i.e., step cycle) and a broader integration window incorporating the opponent motion of the two feet (i.e., gait cycle). A recent study suggests that neural tracking of these nested kinematic structures contributes to the spatiotemporal integration of visual BM cues in different manners (Shen, Lu, Yuan, et al., 2023). However, it remains open whether and how the cortical tracking of hierarchical rhythmic structures underpins the AVI of BM information.

To tackle this issue, we recorded electroencephalogram (EEG) signals from participants who viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions in Experiments 1a & 1b (Fig. 1). An enhanced cortical tracking effect in the AV condition compared to each unisensory condition will indicate significant multisensory gains. Moreover, we adopted an additive model to classify multisensory integration based on the AV vs A+V comparison. This model assumes independence between inputs from each sensory modality and distinguishes among sub-additive (AV < A+V), additive (AV = A+V), and super-additive (AV > A+V) response modes (see a review by Stevenson et al., 2014). The additive mode represents a linear combination between two modalities. In contrast, the super-additive and sub-additive modes indicate non-linear interaction processing, either with potentiated neural activation to facilitate the perception or detection of near-threshold signals (super-additive) or a deactivation mechanism to minimize the processing of redundant information cross-modally (sub-additive) (Laurienti et al., 2005; Metzger et al., 2020; Stanford et al., 2005; Wright et al., 2003). Distinguishing among these integration modes may help elucidate the neural mechanism underlying AVI in specific contexts.

Illustrations of audiovisual stimuli and experimental procedures.

The illustration was based on stimuli with a gait-cycle frequency of 1 Hz. (a) Visual stimuli. The left panel depicts the static schematic of upright and inverted point-light walkers. The right panel shows the keyframes from a gait cycle of the BM sequence. The colors of dots and lines between dots are for illustration only and are not shown in the experiments. (b) Auditory stimuli. The auditory sequences contain periodic impulses of footstep sounds whose peak amplitudes occur around the points when the foot strikes the ground. The duration of two successive impulses defines the gait cycle of footstep sounds, which is temporally congruent (Con) or incongruent (InC) with the visual stimuli. (c) Experimental procedure and design. The color of the visual stimuli changed one or two times within 6 s in the catch trials but did not change in the experimental trials. Participants were required to report the number of changes when the point-light stimulus was replaced by a red fixation. In Experiment 1, participants viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions. The visual stimulus was the BM sequence in the V and AV conditions but a static frame from the sequence in the A condition. Experiment 2 included only the AV condition with different stimulus orientations (upright vs. inverted) and audiovisual congruency (congruent vs. incongruent).

Experiment 2 examined to what extent the AVI effect was specific to the multisensory processing of BM by using non-BM (inverted visual stimuli) as a control. Inversion disrupts the unique, gravity-compatible kinematic features of BM but not the rhythmic signals generated by low-level motion cues (Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Wang et al., 2022), thus is expected to interfere with the BM-specific neural processing. Participants perceived visual BM stimuli accompanied by temporally congruent or incongruent BM sounds.

Comparing the congruency effect in neural responses between the upright and inverted conditions allowed us to verify whether the AVI of BM involves a mechanism distinct from that underlies the AVI of non-BM. In addition, we further explored the possible linkage between the BM-specific neural tracking effect and observers’ autistic traits. Previous behavioral studies found reduced orienting to audiovisually synchronized BM stimuli in ASD (Falck-Ytter et al., 2018). Since individuals with varying social cognitive abilities lie on a continuum extending from clinical to nonclinical populations with different levels of autistic traits, we investigated whether cortical tracking of audiovisual BM correlates with individuals’ autistic traits, as measured by the Autism-Spectrum Quotient (AQ) (Baron-Cohen et al., 2001).

Results

In all experiments, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed one or two times throughout the trial, and there was no color change in other trials. Participants were required to detect the color change of visual stimuli (0-2 times during one trial) to maintain attention. Behavioral analysis on all trials showed that their performances for the task were generally high and equally well in all conditions of Experiment 1a (mean accuracy > 98%; F (2, 46) = 0.814, p = .450, ƞp2 = 0.034), Experiment 1b (mean accuracy > 98%; F (2, 46) = 0.615, p = .545, ƞp2 = 0.026), and Experiment 2 (mean accuracy > 98%; F (3, 69) = 0.493, p = .688, ƞp2 = 0.021), indicating comparable attention state across conditions. The catch trials were excluded from the following EEG analysis.

Cortical tracking of rhythmic structures in audiovisual BM reveals AVI

Experiment 1a

In Experiment 1a, we examined the cortical tracking of rhythmic BM information under V, A, and AV conditions (Fig. 1c). We were interested in two critical rhythmic structures in the walking motion sequence, i.e., the gait cycle and the step cycle (Fig. 1a and b). During walking, each step of the left or right foot occurs alternatively to form a step cycle, and the antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, the frequency of a full gait cycle is 1 Hz, and the step-cycle frequency is 2 Hz. The strength of the cortical tracking effect was quantified by the amplitude peaks emerging from the EEG spectra at these frequencies.

As shown in the grand average amplitude spectra (Fig. 2a), both the responses in three conditions showed clear peaks at step-cycle frequency (2 Hz; V: t (23) = 6.963, p < 0.001; A: t (23) = 6.073, p < .001; AV: t (23) = 7.054, p < 0.001; FDR corrected). In contrast, at gait-cycle frequency (1 Hz), only the response to AV stimulation showed significant peaks (V: t (23) = -2.072, p = 0.975; A: t (23) = -0.054, p = 0.521; AV: t (23) = 4.059, p < 0.001; FDR corrected). Besides, we also observed significant peaks at 4 Hz in all three conditions (ps < 0.001, FDR corrected), which showed a similar audiovisual integration mode as 2 Hz (see more details in Supplementary Information).

Cortical tracking of visual (V), auditory (A), and audiovisual (AV) BM signals at gait-cycle and step-cycle frequencies.

(a) & (d) The amplitude spectra of EEG response in three conditions in Experiment 1a and Experiment 1b, respectively. The solid lines show the grand average amplitude over all electrodes and subjects. The shaded regions depict standard errors of the group mean. Asterisks indicate significant spectra peaks (one-sample t-test against zero; p < .05, FDR corrected). (b) & (e) The normalized amplitude at gait-cycle frequency in the AV condition exceeded the arithmetical sum of those in V and A conditions (AV > A+V), (c) & (f) but the normalized amplitude at step-cycle frequency in the AV condition was comparable to the sum of V and A (AV = A+V). Colored dots represent individual data in each condition. Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; m.s.: .05< p < .10; n.s.: p > .05.

Furthermore, we directly compared the cortical tracking effects between different conditions via a two-tailed paired t-test. At both 1 Hz (Fig. 2b) and 2 Hz (Fig. 2c), the amplitude in the AV condition was greater than that in the V condition (1 Hz: t (23) = 4.664, p < 0.001, Cohen’s d = 0.952; 2 Hz: t (23) = 5.132, p < 0.001, Cohen’s d = 1.048) and the A condition (1 Hz: t (23) = 2.391, p = 0.025, Cohen’s d = 0.488; 2 Hz: t (23) = 3.808, p < 0.001, Cohen’s d = 0.777), respectively, suggesting multisensory gains. More importantly, at 1Hz, the amplitude in the AV condition was significantly larger than the algebraic sum of those in the A and V conditions (t (23) = 3.028, p = 0.006, Cohen’s d = 0.618), indicating a super-additive audiovisual integration effect. While at 2Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.623, p = 0.539, Cohen’s d = -0.127), indicating additive audiovisual integration.

Experiment 1b

To further test whether such cortical tracking effect can apply to stimuli with a different speed, Experiment 1b altered the frequencies of the gait cycle and the corresponding step cycle to 0.83 Hz and 1.67 Hz while adopting the same paradigm as Experiment 1a. Consistent with Experiment 1a, the frequency-domain analysis revealed significant cortical tracking of the audiovisual stimuli at the new speeds. As shown in Fig. 2d, both the responses to V, A, and AV stimuli showed clear peaks at step-cycle frequency (1.67 Hz; V: t (23) = 3.473, p = .001; A: t (23) = 9.194, p < .001; AV: t (23) = 8.756, p < .001; FDR corrected) and its harmonics (3.33 Hz, ps < .001, FDR corrected). In contrast, at gait-cycle frequency (0.83 Hz), only the response to AV stimuli showed significant peaks (V: t (23) = -1.125, p = .846; A: t (23) = -2.449, p = .989; AV: t (23) = 3.052, p = .003; FDR corrected).

At both 0.83 Hz (Fig. 2e) and 1.67 Hz (Fig. 2f), the amplitude in the AV condition was stronger or marginally stronger than that in the V condition (0.83 Hz: t (23) = 2.665, p = .014, Cohen’s d = 0.544; 1.67 Hz: t (23) = 6.380, p < .001, Cohen’s d = 1.302) and the A condition (0.83 Hz: t (23) = 3.625, p < .001, Cohen’s d = 0.740; 1.67 Hz: t (23) = 1.752, p = .093, Cohen’s d = 0.358), respectively, suggesting multisensory gains. More importantly, at 0.83 Hz, the amplitude in the AV condition was significantly larger than the sum of those in the A and V conditions (t (23) = 3.240, p = .004, Cohen’s d = 0.661), indicating a super-additive audiovisual integration effect. While at 1.67 Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.735, p = .470, Cohen’s d = -0.150), indicating linear audiovisual integration. Significant peaks were also observed at 3.33 Hz in all three conditions (ps < 0.001, FDR corrected), which showed similar audiovisual integration mode as 1.67 Hz (see more details in Supplementary Information).

In summary, results from Experiments 1a & 1b consistently showed that the cortical tracking of the audiovisual signals at different temporal scales exhibit distinct audiovisual integration modes, i.e., the super-additive effect at gait-cycle frequency and the additive effect at step-cycle frequency, indicating that the cortical tracking effects at the two temporal scales might be driven by functionally dissociable mechanisms.

Cortical tracking of higher-order rhythmic structure contributes to the AVI of BM

To further explore whether and how the cortical tracking of rhythmic structures contributes to the specialized audiovisual processing of BM, both upright and inverted BM stimuli were adopted in Experiment 2. The task and the frequencies of visual stimuli in Experiment 2 were same as Experiment 1a. Specifically, participants were required to perform the change detection task when perceiving upright and inverted visual BM sequences (1 Hz for gait-cycle frequency and 2 Hz for step-cycle frequency) accompanied by frequency congruent (1 Hz) or incongruent (0.6 Hz and 1.4 Hz) footstep sounds (Fig. 1c). The audiovisual congruency effect, characterized by stronger neural responses in the audiovisual congruent condition compared with the incongruent condition, can be taken as an index of AVI (Fleming et al., 2020; Maddox et al., 2015; Wuerger, Crocker-Buque, et al., 2012). A stronger congruency effect in the upright condition relative to the inverted condition characterizes an AVI process specific to BM information.

We contrasted the congruency effect between the upright and inverted conditions to search for clusters showing a significant difference, which equaled identifying an interaction effect, using a cluster-based permutation test over all electrodes (n = 1000, alpha = 0.05; see Methods). At 1 Hz, the congruency effect in the upright condition was significantly stronger than that in the inverted condition in a cluster at the right hemisphere (Fig. 3a, lower panel, p = 0.029; C2, CPz, CP2, CP4, CP6, Pz, P2, P4, P6). Then, we averaged the amplitude of electrodes within the significant cluster and performed two-tailed paired t-tests to examine whether the congruency was significant in the upright and the inverted conditions, respectively. Results showed that (Fig. 3b) audiovisual congruency enhanced the oscillatory amplitude only for upright BM (t (23) = 4.632, p < 0.001, Cohen’s d = 0.945) but not when visual BM was inverted (t (23) = 0.480, p = 0.635, Cohen’s d = 0.098). Together, these findings suggest that cortical tracking of the high-order gait cycles involves a domain-specific process exclusively engaged in the AVI of BM.

Cortical tracking at gait-cycle rather than step-cycle frequency contributes to the BM-specific audiovisual integration (AVI) effect.

The lower panels in (a) and (d) depict the topographic maps of the BM-specific AVI effect, measured by the difference of congruency effect between the Upright and Inverted conditions at 1 Hz and 2 Hz, respectively. A significantly enhanced congruency effect in the upright condition relative to the inverted condition was observed at 1 Hz (marked by black dots) but not at 2 Hz. The amplitude at these significant electrodes was averaged to quantify the congruency effect for the upright and inverted conditions at 1 Hz (b) and 2 Hz (e). Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; n.s.: p > .05. Individuals’ autistic traits correlated with the BM-specific AVI effect at 1 Hz (c) but not at 2 Hz (f). Shaded regions indicate the 95% confidence intervals.

In contrast, at 2 Hz, no cluster showed a significantly different congruency effect between the upright and inverted conditions (Fig. 3d). We then conducted further analysis based on the electrodes yielded by 1 Hz as marked in Fig. 3a. Results showed that both upright and inverted stimuli induced a significant congruency effect at 2 Hz (Fig. 3e; Upright: t (23) = 3.096, p = 0.005, Cohen’s d = 0.632; Inverted: t (23) = 2.672, p = 0.014, Cohen’s d = 0.545). These findings suggest that neural tracking of the lower-order step cycles is associated with a domain-general AVI process mostly driven by temporal correspondence in physical stimuli.

To verify the dissociation of the BM-specific AVI effects between the gait-cycle frequency and the step-cycle frequencies, we performed a three-way repeated-measures ANOVA with cycle frequency (1 Hz vs. 2 Hz), orientation (upright vs. inverted), and audiovisual congruency (congruent vs. incongruent) as within-subject factors. We converted the amplitude data at each frequency condition into Z-scores to reduce the potential influence of amplitude magnitude difference between the two frequencies. The analysis of these data revealed a significant three-way interaction (F (1,23) = 7.501, p = 0.012, ƞp2 = 0.246), further supporting that the audiovisual integration processing of BM is different between 1 Hz and 2 Hz.

BM-specific cortical tracking correlates with autistic traits

Furthermore, we examined the link between individuals’ autistic traits and the neural responses underpinning the AVI of BM, measured by the difference of congruency effect between the upright and the inverted BM conditions, using Pearson correlation analysis. After removing one outlier (whose neural response exceeded 3 SD from the group mean), we observed an evident negative correlation between individuals’ AQ scores and their neural responses at 1 Hz (Fig. 3c, r = -0.493, p = 0.017) but not at 2 Hz (Fig. 3f, r = -0.158, p = .460). The lack of significant results at 2 Hz was not attributable to electrode selection bias based on the significant cluster at 1 Hz, as similar results were observed when we performed the analysis on clusters showing non-selective significant congruency effects at 2 Hz (see the control analysis in Supplementary Information for details). Besides, we split the participants based on their median AQ score and found that, compared with the high AQ group, the low AQ group showed a greater BM-specific cortical tracking effect at 1 Hz but not at 2 Hz. These findings provide further support to the possible linkage between social cognition and cortical tracking of biological motion as well as its dissociation at the two temporal scales (see more details in Supplementary Information).

Discussion

The current study investigated the neural implementation for the AVI of human BM information. We found that, even under a motion-irrelevant color detection task, observers’ neural activity tracked the temporally corresponding audiovisual BM signals at the frequencies of two rhythmic structures, i.e., the higher-order structure of gait cycle at a larger integration window and the basic-level structure of step cycle at a smaller integration window. The strength of these cortical tracking effects was enhanced under the audiovisual condition than in the visual-only or auditory-only condition, indicating multisensory gains. More crucially, although the cortical tracking of both gait-cycle and step-cycle gain benefits from multisensory correspondence, the mechanisms underlying these two processes appear to be different. At step-cycle frequency, the cortical tracking effect in the AV condition equaled the additive sum of the unisensory conditions. Such linear integration may result from concurrent, independent processing of unisensory inputs without additional interaction of them (Stein et al., 2009). In contrast, at gait-cycle frequency, the congruent audiovisual signals led to a super-additive multisensory enhancement over the linear combination of auditory and visual conditions (AV > A+V), despite that there was no evident cortical tracking effect in the visual condition, different from previous findings obtained with a motion-relevant change detection task (Shen, Lu, Yuan, et al., 2023). This super-additive multisensory enhancement may bring about decreased thresholds of detection and identification (Stanford et al., 2005), allowing us to achieve a more clear and stable perception of the external environment and detect weak stimulus changes in time and respond adaptively.

Furthermore, results from Experiment 2 demonstrated that the cortical tracking of rhythmic structure corresponding to the gait cycle rather than the step cycle underlies the specialized processing of audiovisual BM information. In particular, the AVI effect at step-cycle frequency was significant for both upright and inverted BM signals and comparable between the two conditions, while the AVI effect at gait-cycle frequency was only significant in the upright condition and was greater than that in the inverted condition. The inversion effect has long been regarded as a marker of the specificity of BM processing in numerous behavioral and neuroimaging studies (Grossman & Blake, 2001; Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Vallortigara & Regolin, 2006; Wang et al., 2014; Wang & Jiang, 2012; Wang et al., 2022). Our current findings of the inversion effect in the cortical tracking of audiovisual BM at the gait-cycle frequency suggest that the neural encoding of the higher-order rhythmic structure reflects the AVI of BM and contributes to the specialized processing of BM information. In contrast, the cortical tracking of the step-cycle may reflect the integration of basic motion signals and corresponding sounds. Together, these results reveal that the neural tracking of rhythmic structures at different temporal scales plays distinct roles in the AVI of BM, which may result from the interplay of stimulus-driven and domain-specific mechanisms.

Besides the temporal dynamics of neural activity revealed by the cortical tracking process, we found that the BM-specific AVI effect was associated with neural activity in the right temporoparietal electrodes. This finding likely relates to the activation of the right posterior superior temporal sulcus (pSTS), a region responding to both auditory and visual BM information and being causally involved in BM perception (Bidet-Caulet et al., 2005; Grossman et al., 2005; Wang et al., 2022). While previous fMRI studies have observed STS activation when processing spatial or semantic correspondence between audiovisual BM (Meyer et al., 2011; Wuerger, Parkes, et al., 2012), whether this region also engages in the audiovisual processing of BM signals based on temporal correspondence remains unknown. The current study provides preliminary evidence for such a possibility, inviting future research to localize the exact source of the multisensory integration processes based on imaging data with high spatial and temporal resolutions, such as MEG.

Cortical tracking of external rhythms is also described as cortical entrainment in a broad sense (Ding et al., 2016; Obleser & Kayser, 2019). Controversy remains regarding the involvements of endogenous neural oscillations and stimulus-evoked responses in these processes (Duecker et al., 2024), as it is challenging to fully dissociate these components due to their intricate interplay (Herrmann et al., 2016; Hosseinian et al., 2021). Despite the complexity of the neuronal mechanisms, previous research suggests that cortical tracking or entrainment plays a role in the multisensory processing of simple or discrete rhythmic signals (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). These findings may partially explain the non-selective AVI effect at the step-cycle in the current study. However, we found that the cortical tracking of the higher-order rhythmic structure formed by spatiotemporal integration of meaningful BM information (i.e., the gait cycle of upright walkers rather than inverted walkers) is selectively engaged in the AVI of BM, suggesting that the multisensory processing of natural continuous stimuli may involve unique mechanisms besides the purely stimulus-driven AVI process. These findings advance our understanding of the recently proposed view that multi-timescale neural processes coordinate multisensory integration (Senkowski & Engel, 2024), especially from the perspective of natural stimuli processing. Similar to BM, other natural rhythmic stimuli, like speech and music, also convey hierarchical structures that can entrain neural oscillations at different temporal scales, both in unisensory (Ding et al., 2016; Doelling & Poeppel, 2015) and multisensory contexts (Biau et al., 2022; Crosse et al., 2015; Nozaradan et al., 2016). Possibly, the audiovisual processing of these stimuli engages multi-scale neural coding mechanisms that play distinct functions in perception. Investigating this issue and comparing the results with BM studies will help complete the picture of how the human brain integrates complex, rhythmic information sampled from different sensory modalities to orchestrate perception in natural scenarios.

Since the scope of the current study mainly focused on the neural processing of BM, the task we employed was unrelated to audiovisual correspondence and not for establishing a direct link between neural responses and behavior. This is a limitation of the current study. A recent study demonstrated that listening to frequency-congruent footstep sounds, compared with incongruent sounds, enhanced the visual search for human walkers but not for non-BM stimuli containing the same rhythmic signals, indicating that audiovisual correspondence specifically enhances the perceptual and attentional processing of BM (Shen, Lu, Wang, et al., 2023). Future research could examine whether the cortical tracking of rhythmic structures plays a functional role in such behaviorally relevant tasks. They could also apply advanced neuromodulation techniques to elucidate the causal relevance of the cortical tracking effect to BM perception (e.g., Kösem et al., 2018, 2020).

Last but not least, our study demonstrated that the selective cortical tracking of higher-level rhythmic structure in audiovisually congruent BM signals negatively correlated with individual autistic traits. This finding highlights the critical role of the neural tracking of audiovisual BM signals in social cognition. It also offers the first evidence that differences in audiovisual BM processing are already present in nonclinical individuals at the neural level and associated with their autistic traits, extending previous behavioral evidence for atypical audiovisual BM processing in ASD populations (Falck-Ytter et al., 2013, 2018; Klin et al., 2009). Meanwhile, given that impaired audiovisual BM processing at the early stage may influence social development and result in cascading consequences for lifetime impairments in social interaction (Falck-Ytter et al., 2018; Klin et al., 2005), it is worth exploring neural tracking of audiovisual BM signals in children, which may pave the way for utilizing it as a potential early neural marker for ASD.

Materials and Methods

Participants

Seventy-two participants (mean age ± SD = 22.4 ± 2.6 years, 35 females) took part in the study, 24 for each of Experiment 1a, Experiment 1b, and Experiment 2. All of them had normal or corrected-to-normal vision and reported no history of neurological, psychiatric, or hearing disorders. They were naïve to the purpose of the study and gave informed consent according to procedures and protocols approved by the institutional review board of the Institute of Psychology, Chinese Academy of Sciences.

Stimuli

Visual stimuli

The visual stimuli (Fig. 1a, left panel) consisted of 13 point-light dots attached to the head and major joints of a human walker (Vanrie & Verfaillie, 2004). The point-light walker was presented at the center of the screen without translational motion. It conveys rhythmic structures specified by recurrent forward motions of bilateral limbs (Fig. 1a, right panel). Each step, regardless of left or right foot, occurs recurrently to form a step cycle. The antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, a full gait cycle took 1 second and was repeated 6 times to form a 6-second walking sequence. That is, the gait-cycle frequency is 1 Hz and the step-cycle frequency is 2 Hz. In Experiment 1b, the gait-cycle frequency was 0.83 Hz and the step-cycle frequency was 1.67 Hz. The gait cycle was repeated 6 times to form a 7.2-second walking sequence.

The stimuli in Experiment 2 were the same as that in Experiment 1a. Meanwhile, the point-light BM was flipped vertically to generate inverted BM (Fig. 1a, left panel), which preserves the temporal structure of the stimuli but distorts its distinctive kinematic features, such as movement that is compatible with the effect of gravity (Shen, Lu, Yuan, et al., 2023; Troje & Westhoff, 2006; Wang et al., 2022).

Auditory stimuli

Auditory stimuli were continuous footstep sounds (6 s) with a sampling rate of 44,100 Hz. As shown in Fig. 1b, in Experiments 1a & 2, the gait-cycle frequency of congruent sounds was 1 Hz, which consisted of two steps or two impulses generated by each foot striking the ground within one gait cycle. The incongruent sounds included a faster (1.4 Hz) and a slower (0.60 Hz) sound. Both congruent and incongruent sounds were generated by manipulating the temporal interval between two successive impulses based on the same auditory stimuli. In Experiment 1b, the gait-cycle frequency of sound was 0.83 Hz.

Stimuli presentation

The visual stimuli were rendered white against a grey background and displayed on a CRT (cathode ray tube) monitor. Participants sat 60 cm from the computer screen (1280×1024 at 60 Hz; High: 37.5 cm; Width: 30 cm), with their heads held stationary on a chinrest. The auditory stimuli were presented binaurally over insert earphones. All stimuli were generated and presented using MATLAB together with the Psychophysics Toolbox (Brainard, 1997; Pelli, 1997).

Procedure and task

Experiment 1a

The experiment was conducted in an acoustically dampened and electromagnetically shielded chamber. Participants completed the task under three conditions (Visual: V; Auditory: A; Audiovisual: AV) with the same procedure (Fig. 1c) except for the stimuli. In the V condition, each trial began with a white fixation cross (0.42° × 0.42°) displayed at the center of a gray background for a random duration (0.8 s to 1 s). Subsequently, a 6-s point-light walker (3.05°×5.47°) walked toward the left or right at a constant walking cycle frequency (1 Hz). To maintain observers’ attention, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed (the RGB values changed from [255 255 255] to [207 207 207]) one or two times throughout the trial. Each change lasted 0.5 s. Observers were required to report the number of changes (0, 1, or 2) via keypresses as accurately as possible after the point-light display was replaced by a red fixation. The next trial started 2–3 s after the response. In the A condition, the 6 s-stimuli were replaced by a visually static BM figure accompanied by continuous footstep sounds. The frequency of footstep sounds was congruent with the frequency of visual BM in the V condition. In the AV condition, the stimuli were temporally congruent visual BM sequences (as in the V condition) and footstep sounds (as in the A condition). Three conditions were conducted in separate blocks. V condition was performed in the middle of A and AV conditions. The order of A and AV conditions was counterbalanced across participants. Each participant completed 40 experimental trials without changes and 10-15 catch trials in each condition, resulting in a total of 150-165 trials. In each condition, participants completed a practice session with 3 trials to get familiar with the task before the formal EEG experiment.

Experiment 1b

The procedure of Experiment 1b was the same as that for Experiment 1a but with two exceptions. First, to test if the cortical tracking effect can apply to stimuli with a different speed, we altered the frequencies of gait and step cycles to 0.83 Hz and 1.67 Hz. Second, we presented the 3 conditions (V, A, and AV) in a completely random order to eliminate the influence of presentation order. To minimize the potential influence of condition switch, we increased the trial number in the practice session from 3 to 14 for each condition.

Experiment 2

The procedure in Experiment 2 was similar to the AV condition in Experiment 1a, except that the visually displayed BM was accompanied by frequency congruent (1 Hz) or incongruent (0.6 or 1.4 Hz) footstep sounds. Each participant completed a total of 76 experiment trials, consisting of 36 congruent-trials, 20 incongruent-trials with faster sounds (1.4 Hz), and 20 incongruent-trials with slower sounds (0.6 Hz). These trials were assigned to 3 blocks based on the frequency of the footstep sounds, with the order of the three frequencies balanced across participants. Besides, an inverted BM was used as a control to investigate whether there is a specialized mechanism tuned to the AVI of life motion signals. The order of upright and inverted conditions was balanced across participants. Meanwhile, we measured the participants’ autistic traits by using the Autism-Spectrum Quotient, or AQ questionnaire (Baron-Cohen et al., 2001). Higher AQ scores indicate a higher level of autistic traits.

EEG recording and analysis

EEG was recorded at 1000 Hz using a SynAmps2 NeuroScan amplifier System with 64 electrodes placed on the scalp according to the international 10-20 system.

Horizontal and vertical eye movements were measured via four additional electrodes placed on the outer canthus of each eye and the inferior and superior areas of the left orbit. Impedances were kept below 5 kΩ for all electrodes.

Preprocessing

The catch trials were excluded from EEG analysis. All preprocessing and further analyses were performed using the FieldTrip toolbox (Oostenveld et al., 2011; http://fieldtriptoolbox.org) in the MATLAB environment. EEG recordings were pass-filtered between 0.1 and 30 Hz, and down-sampled to 100 Hz. Then the continuous EEG data were cut into epochs ranging from -1s to 6 gait cycles (7.2 s in Experiment 1b and 6 s in other experiments) time-locked to the onset of the visual point-light stimuli. The epochs were visually inspected, and trials contaminated with excessive noise were excluded from the analysis. After the trial rejection, eye and cardiac artifacts were removed via independent component analysis based on the Runica algorithm (Bell & Sejnowski, 1995; Jung et al., 2000; Makeig, 2002). Then the cleaned data were re-referenced to the average mastoids (M1 and M2). To minimize the influence of stimulus-onset evoked activity on EEG spectral decomposition, the EEG recording before the onset of the stimulus and the first cycle (1 s in Experiments 1a & 2; 1.2 s in Experiment 1b) of each trial was excluded (Nozaradan et al., 2012a). After that, the EEG epochs were averaged across trials for each participant and condition.

Frequency-Domain analysis and statistics

A Fast Fourier Transform (FFT) with zero padding (1200) was used to convert the averaged EEG signals from the temporal domain to the spectral domain, resulting in a frequency resolution of 0.083 Hz, i.e., 1/12 Hz, which is sufficient for observing neural responses around the frequency of the rhythmic BM structures in all experiments. When performing FFT, a Hanning window was adopted to minimize spectral leakage. Then, to remove the 1/f trend of the response amplitude spectrum and identify spectral peaks, the response amplitude at each frequency was normalized by subtracting the average amplitude measured at the neighboring frequency bins (two bins on each side) (Nozaradan et al., 2012a). We calculated the normalized amplitude separately for each electrode (except for electrooculogram electrodes, CB1, and CB2), participant, and condition.

In Experiment 1, the normalized amplitude in all electrodes was averaged and a right-tailed one-sample t-test against zero was performed on the grand average amplitude to test whether the neural response in each frequency bin showed a significant tracking effect or spectral peak. This test was applied to all frequency bins below 5.33 Hz and multiple comparisons were controlled by false discovery rate (FDR) correction at p < 0.05 (Benjamini & Hochberg, 1995). In Experiment 2, to further identify the BM-specific AVI process, the audiovisual congruency effect was compared between the upright and inverted conditions using a cluster-based permutation test over all electrodes (1000 iterations, requiring a cluster size of at least 2 significant neighbors, a two-sided t-test at p < 0.05 on the clustered data) (Oostenveld et al., 2011; http://fieldtriptoolbox.org). This allowed us to identify the spatial distribution of the BM-specific congruency effect.

eLife Assessment

Wang et al. presented visual (dot) motion and/or the sound of a walking person and found solid evidence that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm, with some demonstration that the gait rhythm is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition). The valuable findings will be of wide interest to those examining biological motion perception and oscillatory processes more broadly. Some of the theoretical interpretations concerning entrainment must remain speculative when the authors cannot dissociate evoked responses from entrained oscillatory effects.

Reviewer #1 (Public review):

Summary:

Shen et al. conducted three experiments to study the cortical tracking of the natural rhythms involved in biological motion (BM), and whether these involve audiovisual integration (AVI). They presented participants with visual (dot) motion and/or the sound of a walking person. They found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm. The gait rhythm specifically is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition, Experiments 1a/b), which is independent of the specific step frequency (Experiment 1b). Furthermore, audiovisual integration during tracking of gait was specific to BM, as it was absent (that is, the audiovisual congruency effect) when the walking dot motion was vertically inverted (Experiment 2). Finally, the study shows that an individual's autistic traits are negatively correlated with the BM-AVI congruency effect.

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

Weaknesses:

On revision, the authors are careful not to overinterpret an analysis where the statistical test is not independent from the data (channel) selection criterion.

Reviewer #2 (Public review):

Summary:

The authors evaluate spectral changes in electroencephalography (EEG) data as a function of the congruency of audio and visual information associated with biological motion (BM) or non-biological motion. The results show supra-additive power gains in the neural response to gait dynamics, with trials in which audio and visual information was presented simultaneously producing higher average amplitude than the combined average power for auditory and visual conditions alone. Further analyses suggest that such supra-additivity is specific to BM and emerges from temporoparietal areas. The authors also find that the BM-specific supra-additivity is negatively correlated with autism traits.

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

In the first revisions of the paper, the manuscript better relays the results and anticipates analyses, and this version adequately resolves some concerns I had about analysis details. In a further revision, it is clarified better how the results relate to the various competing hypotheses on how biological motion is processed.

Weaknesses:

Still, it is my view that the findings of the study are basic neural correlate results that offer only minimal constraint towards the question of how the brain realizes the integration of multisensory information in the service of biological motion perception, and the data do not address the causal relevance of observed neural effects towards behavior and cognition. The presence of an inversion effect suggests that the supra-additivity is related to cognition, but that leaves open whether any detected neural pattern is actually consequential for multi-sensory integration (i.e., correlation is not causation). In other words, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated in the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion that offer some constraint toward mechanism, and it is possible that the effects are behaviorally relevant, but based on the current task and associated analyses this has not been shown (or could not have been, given the paradigm).

Author response:

The following is the authors’ response to the previous reviews.

Public Reviews:

Reviewer #1 (Public review):

Summary:

Shen et al. conducted three experiments to study the cortical tracking of the natural rhythms involved in biological motion (BM), and whether these involve audiovisual integration (AVI). They presented participants with visual (dot) motion and/or the sound of a walking person. They found that EEG activity tracks the step rhythm, as well as the gait (2-step cycle) rhythm. The gait rhythm specifically is tracked superadditively (power for A+V condition is higher than the sum of the A-only and V-only condition,

Experiments 1a/b), which is independent of the specific step frequency (Experiment 1b). Furthermore, audiovisual integration during tracking of gait was specific to BM, as it was absent (that is, the audiovisual congruency effect) when the walking dot motion was vertically inverted (Experiment 2). Finally, the study shows that an individual's autistic traits are negatively correlated with the BM-AVI congruency effect.

Strengths:

The three experiments are well designed and the various conditions are well controlled. The rationale of the study is clear, and the manuscript is pleasant to read. The analysis choices are easy to follow, and mostly appropriate.

Weaknesses:

On revision, the authors are careful not to overinterpret an analysis where the statistical test is not independent from the data (channel) selection criterion.

Thanks for the suggestion and we have done this according to your recommendations below.

Reviewer #1 (Recommendations for the authors):

Re: the double-dipping concern: I appreciate the revision. Just to clarify: my concern rests with the selection of *electrodes* based on the interaction test for the 1Hz condition. The 2Hz condition analogous test yields no significant electrodes. You perform subsequent tests (t-tests and 3-way interaction) on the data averaged across the electrodes that were significant for the 1Hz condition. Therefore, these tests will be biased to find a pattern reflecting an interaction at 1Hz, while no similar bias exists for an effect at 2Hz. Therefore, there is a bias to observe a 3-way interaction, and simple effects compatible with a 2-way interaction only for 1Hz, not for 2Hz (which is exactly what you found). There is no good statistical alternative here, I appreciate that, but the bias exists nonetheless. I think the wording is improved in this revision, and the evidence is convincing even in light of this bias.

We are grateful for your thoughtful comments on the analytical methods. We appreciate your concerns regarding the potential bias of examining 3-way interaction based on electrodes yielding a 2-way interaction effect. To address this issue, we have conducted a bias-free analysis based on electrodes across the whole brain. The results showed a similar pattern of 3-way interaction as previously reported (p = 0.051), suggesting that the previous findings might not be caused by electrode selection. Given that the main results of Experiment 2 were not based on whole-brain analysis, we did not involve this analysis in the main text, and we have removed the three-way interaction results based on selected electrodes from the manuscript to reduce potential concerns. It is also noteworthy that, when performing analyses based on channels independent of the interaction effect at 1 Hz (i.e., significant congruency effects in the upright and inverted conditions, respectively, at 2Hz), we got similar results as reported in the main text (i.e., non-significant interaction and correlation at 2 Hz). These results were presented in the supplementary file in previous versions and mentioned in the correlation part of the Results section (see Fig. S2). Once again, we sincerely appreciate your careful review of our research. We hope the abovementioned points adequately address your concern.

Reviewer #2 (Public review):

Summary:

The authors evaluate spectral changes in electroencephalography (EEG) data as a function of the congruency of audio and visual information associated with biological motion (BM) or non-biological motion. The results show supra-additive power gains in the neural response to gait dynamics, with trials in which audio and visual information was presented simultaneously producing higher average amplitude than the combined average power for auditory and visual conditions alone. Further analyses suggest that such supra-additivity is specific to BM and emerges from temporoparietal areas. The authors also find that the BM-specific supra-additivity is negatively correlated with autism traits.

Strengths:

The manuscript is well-written, with a concise and clear writing style. The visual presentation is largely clear. The study involves multiple experiments with different participant groups. Each experiment involves specific considered changes to the experimental paradigm that both replicate the previous experiment's finding yet extend it in a relevant manner.

In the first revisions of the paper, the manuscript better relays the results and anticipates analyses, and this version adequately resolves some concerns I had about analysis details. In a further revision, it is clarified better how the results relate to the various competing hypotheses on how biological motion is processed.

Weaknesses:

Still, it is my view that the findings of the study are basic neural correlate results that offer only minimal constraint towards the question of how the brain realizes the integration of multisensory information in the service of biological motion perception, and the data do not address the causal relevance of observed neural effects towards behavior and cognition. The presence of an inversion effect suggests that the supraadditivity is related to cognition, but that leaves open whether any detected neural pattern is actually consequential for multi-sensory integration (i.e., correlation is not causation). In other words, the fact that frequency-specific neural responses to the [audio & visual] condition are stronger than those to [audio] and [visual] combined does not mean this has implications for behavioral performance. While the correlation to autism traits could suggest some relation to behavior and is interesting in its own right, this correlation is a highly indirect way of assessing behavioral relevance. It would be helpful to test the relevance of supra-additive cortical tracking on a behavioral task directly related to the processing of biological motion to justify the claim that inputs are being integrated in the service of behavior. Under either framework, cortical tracking or entrainment, the causal relevance of neural findings toward cognition is lacking.

Overall, I believe this study finds neural correlates of biological motion that offer some constraint toward mechanism, and it is possible that the effects are behaviorally relevant, but based on the current task and associated analyses this has not been shown (or could not have been, given the paradigm).

Reviewer #2 (Recommendations for the authors):

Thank you for your revisions; I have updated the Strengths section, and reworded the weaknesses section. I now concede that the neural effects observed offer some constraint towards what the neural mechanisms for AV integration for BM are, whereas in my previous review, I said too strongly that these results do not offer any information about mechanism.

Thank you again for your insightful thoughts and comments on our research. They have contributed greatly to enhancing the discussion of the article and provided valuable inspiration for future exploration of causal mechanisms.

Version 4

Abstract

Abstract

When observing others’ behaviors, we continuously integrate their movements with the corresponding sounds to enhance perception and develop adaptive responses. However, how the human brain integrates these complex audiovisual cues based on their natural temporal correspondence remains unclear. Using electroencephalogram (EEG), we demonstrated that rhythmic cortical activity tracked the hierarchical rhythmic structures in audiovisually congruent human walking movements and footstep sounds. Remarkably, the cortical tracking effects exhibit distinct multisensory integration modes at two temporal scales: an additive mode in a lower-order, narrower temporal integration window (step-cycle) and a super-additive enhancement in a higher-order, broader temporal window (gait-cycle). Furthermore, while neural responses at the lower-order timescale reflect a domain-general audiovisual integration process, cortical tracking at the higher-order timescale is exclusively engaged in the integration of biological motion cues. In addition, only this higher-order, domain-specific cortical tracking effect correlates with individuals’ autistic traits, highlighting its potential as a neural marker for autism spectrum disorder. These findings unveil the multifaceted mechanism whereby rhythmic cortical activity supports the multisensory integration of human motion, shedding light on how neural coding of hierarchical temporal structures orchestrates the processing of complex, natural stimuli across multiple timescales.

Introduction

The perception of biological motion (BM), the movements of living creatures, is a fundamental ability of the human visual system. Extensive evidence shows that humans can readily perceive BM from a visual display depicting just a handful of light dots attached to the head and major joints of a moving person (Blake & Shiffrar, 2007). Nevertheless, in real life, BM perception often occurs in multisensory contexts. For instance, one may simultaneously hear footstep sounds while seeing others walking. The integration of these visual and auditory BM cues facilitates the detection, discrimination, and attentional processing of BM (Mendonça et al., 2011; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2013; van der Zwan et al., 2009). Notably, such benefits are diminished when the visual BM is deprived of characteristic kinematic cues but not low-level motion attributes (Brooks et al., 2007; Shen, Lu, Wang, et al., 2023; Thomas & Shiffrar, 2010), and the temporal windows of perceptual audiovisual synchrony are different between BM and non-BM stimuli (Arrighi et al., 2006; Saygin et al., 2008), highlighting the specificity of audiovisual BM processing. This specificity may relate to the evolutionary significance of BM and its pivotal role in social situations. In particular, integrating multisensory BM cues is foundational for perceiving and attending to other people and developing further social interaction. Such ability is usually compromised in people with social deficits, such as individuals with autism spectrum disorder (ASD) (Falck-Ytter et al., 2013, 2018; Feldman et al., 2018). These findings underline the unique contribution of multisensory BM processing to human perception and social cognition. However, despite the behavioral evidence, the neural coding of audiovisual BM cues and its possible link with individuals’ social cognitive capability remains largely unexplored.

An intrinsic property of human movements (such as walking and running) is that they are rhythmic and accompanied by frequency-congruent sounds. The audiovisual integration (AVI) of such rhythmic stimuli may involve a process that brain activity aligning with and tracking external rhythms, revealed by increased power or phase coherence of neural oscillations at corresponding frequencies (Bauer et al., 2020; Ding et al., 2016; Obleser & Kayser, 2019). Studies based on simple or discrete stimuli show that temporal congruency in auditory and visual rhythms significantly enhances the cortical tracking of rhythmic stimulations in both modalities (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). Unlike these stimuli, BM conveys complex hierarchical rhythmic structures corresponding to integration windows at multiple temporal scales. For example, the human locomotion movement has a narrower integration window consisting of each step (i.e., step cycle) and a broader integration window incorporating the opponent motion of the two feet (i.e., gait cycle). A recent study suggests that neural tracking of these nested kinematic structures contributes to the spatiotemporal integration of visual BM cues in different manners (Shen, Lu, Yuan, et al., 2023). However, it remains open whether and how the cortical tracking of hierarchical rhythmic structures underpins the AVI of BM information.

To tackle this issue, we recorded electroencephalogram (EEG) signals from participants who viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions in Experiments 1a & 1b (Fig. 1). An enhanced cortical tracking effect in the AV condition compared to each unisensory condition will indicate significant multisensory gains. Moreover, we adopted an additive model to classify multisensory integration based on the AV vs A+V comparison. This model assumes independence between inputs from each sensory modality and distinguishes among sub-additive (AV < A+V), additive (AV = A+V), and super-additive (AV > A+V) response modes (see a review by Stevenson et al., 2014). The additive mode represents a linear combination between two modalities. In contrast, the super-additive and sub-additive modes indicate non-linear interaction processing, either with potentiated neural activation to facilitate the perception or detection of near-threshold signals (super-additive) or a deactivation mechanism to minimize the processing of redundant information cross-modally (sub-additive) (Laurienti et al., 2005; Metzger et al., 2020; Stanford et al., 2005; Wright et al., 2003). Distinguishing among these integration modes may help elucidate the neural mechanism underlying AVI in specific contexts.

Illustrations of audiovisual stimuli and experimental procedures.

The illustration was based on stimuli with a gait-cycle frequency of 1 Hz. (a) Visual stimuli. The left panel depicts the static schematic of upright and inverted point-light walkers. The right panel shows the keyframes from a gait cycle of the BM sequence. The colors of dots and lines between dots are for illustration only and are not shown in the experiments. (b) Auditory stimuli. The auditory sequences contain periodic impulses of footstep sounds whose peak amplitudes occur around the points when the foot strikes the ground. The duration of two successive impulses defines the gait cycle of footstep sounds, which is temporally congruent (Con) or incongruent (InC) with the visual stimuli. (c) Experimental procedure and design. The color of the visual stimuli changed one or two times within 6 s in the catch trials but did not change in the experimental trials. Participants were required to report the number of changes when the point-light stimulus was replaced by a red fixation. In Experiment 1, participants viewed rhythmic point-light walkers or/and listened to the corresponding footstep sounds under visual (V), auditory (A), and audiovisual (AV) conditions. The visual stimulus was the BM sequence in the V and AV conditions but a static frame from the sequence in the A condition. Experiment 2 included only the AV condition with different stimulus orientations (upright vs. inverted) and audiovisual congruency (congruent vs. incongruent).

Experiment 2 examined to what extent the AVI effect was specific to the multisensory processing of BM by using non-BM (inverted visual stimuli) as a control. Inversion disrupts the unique, gravity-compatible kinematic features of BM but not the rhythmic signals generated by low-level motion cues (Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Wang et al., 2022), thus is expected to interfere with the BM-specific neural processing. Participants perceived visual BM stimuli accompanied by temporally congruent or incongruent BM sounds.

Comparing the congruency effect in neural responses between the upright and inverted conditions allowed us to verify whether the AVI of BM involves a mechanism distinct from that underlies the AVI of non-BM. In addition, we further explored the possible linkage between the BM-specific neural tracking effect and observers’ autistic traits. Previous behavioral studies found reduced orienting to audiovisually synchronized BM stimuli in ASD (Falck-Ytter et al., 2018). Since individuals with varying social cognitive abilities lie on a continuum extending from clinical to nonclinical populations with different levels of autistic traits, we investigated whether cortical tracking of audiovisual BM correlates with individuals’ autistic traits, as measured by the Autism-Spectrum Quotient (AQ) (Baron-Cohen et al., 2001).

Results

In all experiments, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed one or two times throughout the trial, and there was no color change in other trials. Participants were required to detect the color change of visual stimuli (0-2 times during one trial) to maintain attention. Behavioral analysis on all trials showed that their performances for the task were generally high and equally well in all conditions of Experiment 1a (mean accuracy > 98%; F (2, 46) = 0.814, p = .450, ƞp2 = 0.034), Experiment 1b (mean accuracy > 98%; F (2, 46) = 0.615, p = .545, ƞp2 = 0.026), and Experiment 2 (mean accuracy > 98%; F (3, 69) = 0.493, p = .688, ƞp2 = 0.021), indicating comparable attention state across conditions. The catch trials were excluded from the following EEG analysis.

Cortical tracking of rhythmic structures in audiovisual BM reveals AVI

Experiment 1a

In Experiment 1a, we examined the cortical tracking of rhythmic BM information under V, A, and AV conditions (Fig. 1c). We were interested in two critical rhythmic structures in the walking motion sequence, i.e., the gait cycle and the step cycle (Fig. 1a & 1b). During walking, each step of the left or right foot occurs alternatively to form a step cycle, and the antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, the frequency of a full gait cycle is 1 Hz, and the step-cycle frequency is 2 Hz. The strength of the cortical tracking effect was quantified by the amplitude peaks emerging from the EEG spectra at these frequencies.

As shown in the grand average amplitude spectra (Fig. 2a), both the responses in three conditions showed clear peaks at step-cycle frequency (2 Hz; V: t (23) = 6.963, p < 0.001; A: t (23) = 6.073, p < .001; AV: t (23) = 7.054, p < 0.001; FDR corrected). In contrast, at gait-cycle frequency (1 Hz), only the response to AV stimulation showed significant peaks (V: t (23) = -2.072, p = 0.975; A: t (23) = -0.054, p = 0.521; AV: t (23) = 4.059, p < 0.001; FDR corrected). Besides, we also observed significant peaks at 4 Hz in all three conditions (ps < 0.001, FDR corrected), which showed a similar audiovisual integration mode as 2 Hz (see more details in Supplementary Information).

Cortical tracking of visual (V), auditory (A), and audiovisual (AV) BM signals at gait-cycle and step-cycle frequencies.

(a) & (d) The amplitude spectra of EEG response in three conditions in Experiment 1a and Experiment 1b, respectively. The solid lines show the grand average amplitude over all electrodes and subjects. The shaded regions depict standard errors of the group mean. Asterisks indicate significant spectra peaks (one-sample t-test against zero; p < .05, FDR corrected). (b) & (e) The normalized amplitude at gait-cycle frequency in the AV condition exceeded the arithmetical sum of those in V and A conditions (AV > A+V), (c) & (f) but the normalized amplitude at step-cycle frequency in the AV condition was comparable to the sum of V and A (AV = A+V). Colored dots represent individual data in each condition. Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; m.s.: .05< p < .10; n.s.: p > .05.

Furthermore, we directly compared the cortical tracking effects between different conditions via a two-tailed paired t-test. At both 1 Hz (Fig. 2b) and 2 Hz (Fig. 2c), the amplitude in the AV condition was greater than that in the V condition (1 Hz: t (23) = 4.664, p < 0.001, Cohen’s d = 0.952; 2 Hz: t (23) = 5.132, p < 0.001, Cohen’s d = 1.048) and the A condition (1 Hz: t (23) = 2.391, p = 0.025, Cohen’s d = 0.488; 2 Hz: t (23) = 3.808, p < 0.001, Cohen’s d = 0.777), respectively, suggesting multisensory gains. More importantly, at 1Hz, the amplitude in the AV condition was significantly larger than the algebraic sum of those in the A and V conditions (t (23) = 3.028, p = 0.006, Cohen’s d = 0.618), indicating a super-additive audiovisual integration effect. While at 2Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.623, p = 0.539, Cohen’s d = -0.127), indicating additive audiovisual integration.

Experiment 1b

To further test whether such cortical tracking effect can apply to stimuli with a different speed, Experiment 1b altered the frequencies of the gait cycle and the corresponding step cycle to 0.83 Hz and 1.67 Hz while adopting the same paradigm as Experiment 1a. Consistent with Experiment 1a, the frequency-domain analysis revealed significant cortical tracking of the audiovisual stimuli at the new speeds. As shown in Fig. 2d, both the responses to V, A, and AV stimuli showed clear peaks at step-cycle frequency (1.67 Hz; V: t (23) = 3.473, p = .001; A: t (23) = 9.194, p < .001; AV: t (23) = 8.756, p < .001; FDR corrected) and its harmonics (3.33 Hz, ps < .001, FDR corrected). In contrast, at gait-cycle frequency (0.83 Hz), only the response to AV stimuli showed significant peaks (V: t (23) = -1.125, p = .846; A: t (23) = -2.449, p = .989; AV: t (23) = 3.052, p = .003; FDR corrected).

At both 0.83 Hz (Fig. 2e) and 1.67 Hz (Fig. 2f), the amplitude in the AV condition was stronger or marginally stronger than that in the V condition (0.83 Hz: t (23) = 2.665, p = .014, Cohen’s d = 0.544; 1.67 Hz: t (23) = 6.380, p < .001, Cohen’s d = 1.302) and the A condition (0.83 Hz: t (23) = 3.625, p < .001, Cohen’s d = 0.740; 1.67 Hz: t (23) = 1.752, p = .093, Cohen’s d = 0.358), respectively, suggesting multisensory gains. More importantly, at 0.83 Hz, the amplitude in the AV condition was significantly larger than the sum of those in the A and V conditions (t (23) = 3.240, p = .004, Cohen’s d = 0.661), indicating a super-additive audiovisual integration effect. While at 1.67 Hz, the amplitude in the AV condition was comparable to the unisensory sum (t (23) = -0.735, p = .470, Cohen’s d = -0.150), indicating linear audiovisual integration. Significant peaks were also observed at 3.33 Hz in all three conditions (ps < 0.001, FDR corrected), which showed similar audiovisual integration mode as 1.67 Hz (see more details in Supplementary Information).

In summary, results from Experiments 1a & 1b consistently showed that the cortical tracking of the audiovisual signals at different temporal scales exhibit distinct audiovisual integration modes, i.e., the super-additive effect at gait-cycle frequency and the additive effect at step-cycle frequency, indicating that the cortical tracking effects at the two temporal scales might be driven by functionally dissociable mechanisms.

Cortical tracking of higher-order rhythmic structure contributes to the AVI of BM

To further explore whether and how the cortical tracking of rhythmic structures contributes to the specialized audiovisual processing of BM, both upright and inverted BM stimuli were adopted in Experiment 2. The task and the frequencies of visual stimuli in Experiment 2 were same as Experiment 1a. Specifically, participants were required to perform the change detection task when perceiving upright and inverted visual BM sequences (1 Hz for gait-cycle frequency and 2 Hz for step-cycle frequency) accompanied by frequency congruent (1 Hz) or incongruent (0.6 Hz and 1.4 Hz) footstep sounds (Fig. 1c). The audiovisual congruency effect, characterized by stronger neural responses in the audiovisual congruent condition compared with the incongruent condition, can be taken as an index of AVI (Fleming et al., 2020; Maddox et al., 2015; Wuerger, Crocker-Buque, et al., 2012). A stronger congruency effect in the upright condition relative to the inverted condition characterizes an AVI process specific to BM information.

We contrasted the congruency effect between the upright and inverted conditions to search for clusters showing a significant difference, which equaled identifying an interaction effect, using a cluster-based permutation test over all electrodes (n = 1000, alpha = 0.05; see Methods). At 1 Hz, the congruency effect in the upright condition was significantly stronger than that in the inverted condition in a cluster at the right hemisphere (Fig. 3a, lower panel, p = 0.029; C2, CPz, CP2, CP4, CP6, Pz, P2, P4, P6). Then, we averaged the amplitude of electrodes within the significant cluster and performed two-tailed paired t-tests to examine whether the congruency was significant in the upright and the inverted conditions, respectively. Results showed that (Fig. 3b) audiovisual congruency enhanced the oscillatory amplitude only for upright BM (t (23) = 4.632, p < 0.001, Cohen’s d = 0.945) but not when visual BM was inverted (t (23) = 0.480, p = 0.635, Cohen’s d = 0.098). Together, these findings suggest that cortical tracking of the high-order gait cycles involves a domain-specific process exclusively engaged in the AVI of BM.

Cortical tracking at gait-cycle rather than step-cycle frequency contributes to the BM-specific audiovisual integration (AVI) effect.

The lower panels in (a) and (d) depict the topographic maps of the BM-specific AVI effect, measured by the difference of congruency effect between the Upright and Inverted conditions at 1 Hz and 2 Hz, respectively. A significantly enhanced congruency effect in the upright condition relative to the inverted condition was observed at 1 Hz (marked by black dots) but not at 2 Hz. The amplitude at these significant electrodes was averaged to quantify the congruency effect for the upright and inverted conditions at 1 Hz (b) and 2 Hz (e). Error bars represent ±1 standard error of means. *: p < .05; **: p < .01; ***: p < .001; n.s.: p > .05. Individuals’ autistic traits correlated with the BM-specific AVI effect at 1 Hz (c) but not at 2 Hz (f). Shaded regions indicate the 95% confidence intervals.

In contrast, at 2 Hz, no cluster showed a significantly different congruency effect between the upright and inverted conditions (Fig. 3d). We then conducted further analysis based on the electrodes yielded by 1 Hz as marked in Fig. 3a. Results showed that both upright and inverted stimuli induced a significant congruency effect at 2 Hz (Fig. 3e; Upright: t (23) = 3.096, p = 0.005, Cohen’s d = 0.632; Inverted: t (23) = 2.672, p = 0.014, Cohen’s d = 0.545). These findings suggest that neural tracking of the lower-order step cycles is associated with a domain-general AVI process mostly driven by temporal correspondence in physical stimuli.

BM-specific cortical tracking correlates with autistic traits

Furthermore, we examined the link between individuals’ autistic traits and the neural responses underpinning the AVI of BM, measured by the difference of congruency effect between the upright and the inverted BM conditions, using Pearson correlation analysis. After removing one outlier (whose neural response exceeded 3 SD from the group mean), we observed an evident negative correlation between individuals’ AQ scores and their neural responses at 1 Hz (Fig. 3c, r = -0.493, p = 0.017) but not at 2 Hz (Fig. 3f, r = -0.158, p = .460). The lack of significant results at 2 Hz was not attributable to electrode selection bias based on the significant cluster at 1 Hz, as similar results were observed when we performed the analysis on clusters showing non-selective significant congruency effects at 2 Hz (see the control analysis in Supplementary Information for details). Besides, we split the participants based on their median AQ score and found that, compared with the high AQ group, the low AQ group showed a greater BM-specific cortical tracking effect at 1 Hz but not at 2 Hz. These findings provide further support to the possible linkage between social cognition and cortical tracking of biological motion as well as its dissociation at the two temporal scales (see more details in Supplementary Information).

Discussion

The current study investigated the neural implementation for the AVI of human BM information. We found that, even under a motion-irrelevant color detection task, observers’ neural activity tracked the temporally corresponding audiovisual BM signals at the frequencies of two rhythmic structures, i.e., the higher-order structure of gait cycle at a larger integration window and the basic-level structure of step cycle at a smaller integration window. The strength of these cortical tracking effects was enhanced under the audiovisual condition than in the visual-only or auditory-only condition, indicating multisensory gains. More crucially, although the cortical tracking of both gait-cycle and step-cycle gain benefits from multisensory correspondence, the mechanisms underlying these two processes appear to be different. At step-cycle frequency, the cortical tracking effect in the AV condition equaled the additive sum of the unisensory conditions. Such linear integration may result from concurrent, independent processing of unisensory inputs without additional interaction of them (Stein et al., 2009). In contrast, at gait-cycle frequency, the congruent audiovisual signals led to a super-additive multisensory enhancement over the linear combination of auditory and visual conditions (AV > A+V), despite that there was no evident cortical tracking effect in the visual condition, different from previous findings obtained with a motion-relevant change detection task (Shen, Lu, Yuan, et al., 2023). This super-additive multisensory enhancement may bring about decreased thresholds of detection and identification (Stanford et al., 2005), allowing us to achieve a more clear and stable perception of the external environment and detect weak stimulus changes in time and respond adaptively.

Furthermore, results from Experiment 2 demonstrated that the cortical tracking of rhythmic structure corresponding to the gait cycle rather than the step cycle underlies the specialized processing of audiovisual BM information. In particular, the AVI effect at step-cycle frequency was significant for both upright and inverted BM signals and comparable between the two conditions, while the AVI effect at gait-cycle frequency was only significant in the upright condition and was greater than that in the inverted condition. The inversion effect has long been regarded as a marker of the specificity of BM processing in numerous behavioral and neuroimaging studies (Grossman & Blake, 2001; Ma et al., 2022; Shen, Lu, Yuan, et al., 2023; Simion et al., 2008; Troje & Westhoff, 2006; Vallortigara & Regolin, 2006; Wang et al., 2014; Wang & Jiang, 2012; Wang et al., 2022). Our current findings of the inversion effect in the cortical tracking of audiovisual BM at the gait-cycle frequency suggest that the neural encoding of the higher-order rhythmic structure reflects the AVI of BM and contributes to the specialized processing of BM information. In contrast, the cortical tracking of the step-cycle may reflect the integration of basic motion signals and corresponding sounds. Together, these results reveal that the neural tracking of rhythmic structures at different temporal scales plays distinct roles in the AVI of BM, which may result from the interplay of stimulus-driven and domain-specific mechanisms.

Besides the temporal dynamics of neural activity revealed by the cortical tracking process, we found that the BM-specific AVI effect was associated with neural activity in the right temporoparietal electrodes. This finding likely relates to the activation of the right posterior superior temporal sulcus (pSTS), a region responding to both auditory and visual BM information and being causally involved in BM perception (Bidet-Caulet et al., 2005; Grossman et al., 2005; Wang et al., 2022). While previous fMRI studies have observed STS activation when processing spatial or semantic correspondence between audiovisual BM (Meyer et al., 2011; Wuerger, Parkes, et al., 2012), whether this region also engages in the audiovisual processing of BM signals based on temporal correspondence remains unknown. The current study provides preliminary evidence for such a possibility, inviting future research to localize the exact source of the multisensory integration processes based on imaging data with high spatial and temporal resolutions, such as MEG.

Cortical tracking of external rhythms is also described as cortical entrainment in a broad sense (Ding et al., 2016; Obleser & Kayser, 2019). Controversy remains regarding the involvements of endogenous neural oscillations and stimulus-evoked responses in these processes (Duecker et al., 2024), as it is challenging to fully dissociate these components due to their intricate interplay (Herrmann et al., 2016; Hosseinian et al., 2021). Despite the complexity of the neuronal mechanisms, previous research suggests that cortical tracking or entrainment plays a role in the multisensory processing of simple or discrete rhythmic signals (Covic et al., 2017; Keitel & Müller, 2016; Nozaradan et al., 2012b). These findings may partially explain the non-selective AVI effect at the step-cycle in the current study. However, we found that the cortical tracking of the higher-order rhythmic structure formed by spatiotemporal integration of meaningful BM information (i.e., the gait cycle of upright walkers rather than inverted walkers) is selectively engaged in the AVI of BM, suggesting that the multisensory processing of natural continuous stimuli may involve unique mechanisms besides the purely stimulus-driven AVI process. These findings advance our understanding of the recently proposed view that multi-timescale neural processes coordinate multisensory integration (Senkowski & Engel, 2024), especially from the perspective of natural stimuli processing. Similar to BM, other natural rhythmic stimuli, like speech and music, also convey hierarchical structures that can entrain neural oscillations at different temporal scales, both in unisensory (Ding et al., 2016; Doelling & Poeppel, 2015) and multisensory contexts (Biau et al., 2022; Crosse et al., 2015; Nozaradan et al., 2016). Possibly, the audiovisual processing of these stimuli engages multi-scale neural coding mechanisms that play distinct functions in perception. Investigating this issue and comparing the results with BM studies will help complete the picture of how the human brain integrates complex, rhythmic information sampled from different sensory modalities to orchestrate perception in natural scenarios.

Since the scope of the current study mainly focused on the neural processing of BM, the task we employed was unrelated to audiovisual correspondence and not for establishing a direct link between neural responses and behavior. This is a limitation of the current study. A recent study demonstrated that listening to frequency-congruent footstep sounds, compared with incongruent sounds, enhanced the visual search for human walkers but not for non-BM stimuli containing the same rhythmic signals, indicating that audiovisual correspondence specifically enhances the perceptual and attentional processing of BM (Shen, Lu, Wang, et al., 2023). Future research could examine whether the cortical tracking of rhythmic structures plays a functional role in such behaviorally relevant tasks. They could also apply advanced neuromodulation techniques to elucidate the causal relevance of the cortical tracking effect to BM perception (e.g., Kösem et al., 2018, 2020).

Last but not least, our study demonstrated that the selective cortical tracking of higher-level rhythmic structure in audiovisually congruent BM signals negatively correlated with individual autistic traits. This finding highlights the critical role of the neural tracking of audiovisual BM signals in social cognition. It also offers the first evidence that differences in audiovisual BM processing are already present in nonclinical individuals at the neural level and associated with their autistic traits, extending previous behavioral evidence for atypical audiovisual BM processing in ASD populations (Falck-Ytter et al., 2013, 2018; Klin et al., 2009). Meanwhile, given that impaired audiovisual BM processing at the early stage may influence social development and result in cascading consequences for lifetime impairments in social interaction (Falck-Ytter et al., 2018; Klin et al., 2005), it is worth exploring neural tracking of audiovisual BM signals in children, which may pave the way for utilizing it as a potential early neural marker for ASD.

Materials and Methods

Participants

Seventy-two participants (mean age ± SD = 22.4 ± 2.6 years, 35 females) took part in the study, 24 for each of Experiment 1a, Experiment 1b, and Experiment 2. All of them had normal or corrected-to-normal vision and reported no history of neurological, psychiatric, or hearing disorders. They were naïve to the purpose of the study and gave informed consent according to procedures and protocols approved by the institutional review board of the Institute of Psychology, Chinese Academy of Sciences.

Stimuli

Visual stimuli

The visual stimuli (Fig. 1a, left panel) consisted of 13 point-light dots attached to the head and major joints of a human walker (Vanrie & Verfaillie, 2004). The point-light walker was presented at the center of the screen without translational motion. It conveys rhythmic structures specified by recurrent forward motions of bilateral limbs (Fig. 1a, right panel). Each step, regardless of left or right foot, occurs recurrently to form a step cycle. The antiphase oscillations of limbs during two steps characterize a gait cycle (Shen, Lu, Yuan, et al., 2023). In Experiment 1a, a full gait cycle took 1 second and was repeated 6 times to form a 6-second walking sequence. That is, the gait-cycle frequency is 1 Hz and the step-cycle frequency is 2 Hz. In Experiment 1b, the gait-cycle frequency was 0.83 Hz and the step-cycle frequency was 1.67 Hz. The gait cycle was repeated 6 times to form a 7.2-second walking sequence.

The stimuli in Experiment 2 were the same as that in Experiment 1a. Meanwhile, the point-light BM was flipped vertically to generate inverted BM (Fig. 1a, left panel), which preserves the temporal structure of the stimuli but distorts its distinctive kinematic features, such as movement that is compatible with the effect of gravity (Shen, Lu, Yuan, et al., 2023; Troje & Westhoff, 2006; Wang et al., 2022).

Auditory stimuli

Auditory stimuli were continuous footstep sounds (6 s) with a sampling rate of 44,100 Hz. As shown in Fig. 1b, in Experiments 1a & 2, the gait-cycle frequency of congruent sounds was 1 Hz, which consisted of two steps or two impulses generated by each foot striking the ground within one gait cycle. The incongruent sounds included a faster (1.4 Hz) and a slower (0.60 Hz) sound. Both congruent and incongruent sounds were generated by manipulating the temporal interval between two successive impulses based on the same auditory stimuli. In Experiment 1b, the gait-cycle frequency of sound was 0.83 Hz.

Stimuli presentation

The visual stimuli were rendered white against a grey background and displayed on a CRT (cathode ray tube) monitor. Participants sat 60 cm from the computer screen (1280×1024 at 60 Hz; High: 37.5 cm; Width: 30 cm), with their heads held stationary on a chinrest. The auditory stimuli were presented binaurally over insert earphones. All stimuli were generated and presented using MATLAB together with the Psychophysics Toolbox (Brainard, 1997; Pelli, 1997).

Procedure and task

Experiment 1a

The experiment was conducted in an acoustically dampened and electromagnetically shielded chamber. Participants completed the task under three conditions (Visual: V; Auditory: A; Audiovisual: AV) with the same procedure (Fig. 1c) except for the stimuli. In the V condition, each trial began with a white fixation cross (0.42° × 0.42°) displayed at the center of a gray background for a random duration (0.8 s to 1 s). Subsequently, a 6-s point-light walker (3.05°×5.47°) walked toward the left or right at a constant walking cycle frequency (1 Hz). To maintain observers’ attention, 17%–23% of the trials were randomly selected as catch trials, in which the color of the walker changed (the RGB values changed from [255 255 255] to [207 207 207]) one or two times throughout the trial. Each change lasted 0.5 s. Observers were required to report the number of changes (0, 1, or 2) via keypresses as accurately as possible after the point-light display was replaced by a red fixation. The next trial started 2–3 s after the response. In the A condition, the 6 s-stimuli were replaced by a visually static BM figure accompanied by continuous footstep sounds. The frequency of footstep sounds was congruent with the frequency of visual BM in the V condition. In the AV condition, the stimuli were temporally congruent visual BM sequences (as in the V condition) and footstep sounds (as in the A condition). Three conditions were conducted in separate blocks. V condition was performed in the middle of A and AV conditions. The order of A and AV conditions was counterbalanced across participants. Each participant completed 40 experimental trials without changes and 10-15 catch trials in each condition, resulting in a total of 150-165 trials. In each condition, participants completed a practice session with 3 trials to get familiar with the task before the formal EEG experiment.

Experiment 1b

The procedure of Experiment 1b was the same as that for Experiment 1a but with two exceptions. First, to test if the cortical tracking effect can apply to stimuli with a different speed, we altered the frequencies of gait and step cycles to 0.83 Hz and 1.67 Hz. Second, we presented the 3 conditions (V, A, and AV) in a completely random order to eliminate the influence of presentation order. To minimize the potential influence of condition switch, we increased the trial number in the practice session from 3 to 14 for each condition.

Experiment 2

The procedure in Experiment 2 was similar to the AV condition in Experiment 1a, except that the visually displayed BM was accompanied by frequency congruent (1 Hz) or incongruent (0.6 or 1.4 Hz) footstep sounds. Each participant completed a total of 76 experiment trials, consisting of 36 congruent-trials, 20 incongruent-trials with faster sounds (1.4 Hz), and 20 incongruent-trials with slower sounds (0.6 Hz). These trials were assigned to 3 blocks based on the frequency of the footstep sounds, with the order of the three frequencies balanced across participants. Besides, an inverted BM was used as a control to investigate whether there is a specialized mechanism tuned to the AVI of life motion signals. The order of upright and inverted conditions was balanced across participants. Meanwhile, we measured the participants’ autistic traits by using the Autism-Spectrum Quotient, or AQ questionnaire (Baron-Cohen et al., 2001). Higher AQ scores indicate a higher level of autistic traits.

EEG recording and analysis

EEG was recorded at 1000 Hz using a SynAmps2 NeuroScan amplifier System with 64 electrodes placed on the scalp according to the international 10-20 system.

Horizontal and vertical eye movements were measured via four additional electrodes placed on the outer canthus of each eye and the inferior and superior areas of the left orbit. Impedances were kept below 5 kΩ for all electrodes.

Preprocessing

The catch trials were excluded from EEG analysis. All preprocessing and further analyses were performed using the FieldTrip toolbox (Oostenveld et al., 2011; http://fieldtriptoolbox.org) in the MATLAB environment. EEG recordings were pass-filtered between 0.1 and 30 Hz, and down-sampled to 100 Hz. Then the continuous EEG data were cut into epochs ranging from -1s to 6 gait cycles (7.2 s in Experiment 1b and 6 s in other experiments) time-locked to the onset of the visual point-light stimuli. The epochs were visually inspected, and trials contaminated with excessive noise were excluded from the analysis. After the trial rejection, eye and cardiac artifacts were removed via independent component analysis based on the Runica algorithm (Bell & Sejnowski, 1995; Jung et al., 2000; Makeig, 2002). Then the cleaned data were re-referenced to the average mastoids (M1 and M2). To minimize the influence of stimulus-onset evoked activity on EEG spectral decomposition, the EEG recording before the onset of the stimulus and the first cycle (1 s in Experiments 1a & 2; 1.2 s in Experiment 1b) of each trial was excluded (Nozaradan et al., 2012a). After that, the EEG epochs were averaged across trials for each participant and condition.

Frequency-Domain analysis and statistics

A Fast Fourier Transform (FFT) with zero padding (1200) was used to convert the averaged EEG signals from the temporal domain to the spectral domain, resulting in a frequency resolution of 0.083 Hz, i.e., 1/12 Hz, which is sufficient for observing neural responses around the frequency of the rhythmic BM structures in all experiments. When performing FFT, a Hanning window was adopted to minimize spectral leakage. Then, to remove the 1/f trend of the response amplitude spectrum and identify spectral peaks, the response amplitude at each frequency was normalized by subtracting the average amplitude measured at the neighboring frequency bins (two bins on each side) (Nozaradan et al., 2012a). We calculated the normalized amplitude separately for each electrode (except for electrooculogram electrodes, CB1, and CB2), participant, and condition.

In Experiment 1, the normalized amplitude in all electrodes was averaged and a right-tailed one-sample t-test against zero was performed on the grand average amplitude to test whether the neural response in each frequency bin showed a significant tracking effect or spectral peak. This test was applied to all frequency bins below 5.33 Hz and multiple comparisons were controlled by false discovery rate (FDR) correction at p < 0.05 (Benjamini & Hochberg, 1995). In Experiment 2, to further identify the BM-specific AVI process, the audiovisual congruency effect was compared between the upright and inverted conditions using a cluster-based permutation test over all electrodes (1000 iterations, requiring a cluster size of at least 2 significant neighbors, a two-sided t-test at p < 0.05 on the clustered data) (Oostenveld et al., 2011; http://fieldtriptoolbox.org). This allowed us to identify the spatial distribution of the BM-specific congruency effect.